---
title: "BA_rate_L"
output: html_notebook
author: Karolina Gustavsson
date: 2026-05-14
---

Input data: AMORIS (longitudinal) from Calculating_BA_L.rmd

# 1.0 Preparing

## 1.1 Selection for longitudinal set
```{r}
# Count rows per sampleID
id_counts <- table(to1po2$sampleID)

# Keep only IDs with at least 3 rows
keep_ids <- names(id_counts[id_counts >= 3])

# Subset to create results_df
results_dt <- to1po2[to1po2$sampleID %in% keep_ids, ]
```

## 1.2 Scaling
```{r}
# Define the BioAge variables you standardized
bioage_vars <- c("kdm", "kdm_advance", "phenoage", "phenoage_advance")

# Compute SDs (in years) for each
bioage_sds <- sapply(results_dt[bioage_vars], sd, na.rm = TRUE)

# Create a readable data frame
bioage_scale_df <- data.frame(
  Variable = names(bioage_sds),
  SD_in_years = round(bioage_sds, 2)
)

# Print the result
print(bioage_scale_df)
```

```{r}
write.table(as.data.frame(bioage_scale_df), "bioage_scale_df_table_RMA.csv", sep = ";",
            dec = ",",
            row.names = FALSE,
            quote = FALSE,
            fileEncoding = "Latin1")
```

## 1.2 Load baseline data 
```{r}
SveDem <- read_sas("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/Data_ADRD/SveDem.sas7bdat")
#SveDem

o_amoris <- read_sas("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/Data_ADRD/StudyPop.sas7bdat")
colnames(o_amoris)[colnames(o_amoris) == "Id"] <- "sampleID"
```

## 1.3 Harmonize variables
```{r}
#According to SveDem
o_amoris$S_ADRD <- ifelse(o_amoris$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(o_amoris$sampleID, SveDem$Id)] %in% c(1,2,3,4,5,6,7,8,9), 1, 0) #1-9
o_amoris$S_e_alz <- ifelse(o_amoris$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(o_amoris$sampleID, SveDem$Id)] == 1, 1, 0) #1
o_amoris$S_l_alz <- ifelse(o_amoris$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(o_amoris$sampleID, SveDem$Id)] == 2, 1, 0) #2
o_amoris$S_alz_vas <- ifelse(o_amoris$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(o_amoris$sampleID, SveDem$Id)] == 3, 1, 0) #3
o_amoris$S_vas <- ifelse(o_amoris$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(o_amoris$sampleID, SveDem$Id)] == 4, 1, 0) #4
o_amoris$S_lew <- ifelse(o_amoris$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(o_amoris$sampleID, SveDem$Id)] == 5, 1, 0) #5
o_amoris$S_fro <- ifelse(o_amoris$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(o_amoris$sampleID, SveDem$Id)] == 6, 1, 0) #6
o_amoris$S_par <- ifelse(o_amoris$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(o_amoris$sampleID, SveDem$Id)] == 7, 1, 0) #7
o_amoris$S_uns <- ifelse(o_amoris$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(o_amoris$sampleID, SveDem$Id)] == 8, 1, 0) #8 
o_amoris$S_oth <- ifelse(o_amoris$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(o_amoris$sampleID, SveDem$Id)] == 9, 1, 0) #9  

# Create the new column by matching sampleID from o_amoris to Id in SveDem
o_amoris$S_INDATUM <- SveDem$DIAGNOSIS_DATE_BR[match(o_amoris$sampleID, SveDem$Id)]

# Codes for the ones not in SveDem
alzheimers_codes <- c(
  "G308", "G309", "G301", "F009", "F002", "G300", "F001", "F000",
  "290,1", "304", "290", "290A", "F00", "305", "290B", "G30", "311A"
)

vascular_codes <- c(
  "293", "290E", "F01", "293,1", "F019", "F012", "F018",
  "F013", "F011", "F010", "F0180", "F0183"
)

other_codes <- c(
  "F039", "F23", "F020", "F024", "F028", "F022", "F03-p", "F021", "F023", "F03-P",
  "290,0", "293,9", "293,2", "293,5", "293,3", "293,4", "F0013", "F0010",
  "306", "290X", "F02", "290W", "F03", "294B", "G311",
  "311B", "G318A", "311C", "F051", "311X"
)

all_codes <- unique(c(alzheimers_codes, vascular_codes, other_codes))

#Then we make an overall category
o_amoris$ADRD <- ifelse(
  (o_amoris$Dementia != "" & !is.na(o_amoris$Dementia)) |
    (o_amoris$ATC != "" & !is.na(o_amoris$ATC)) |
    (o_amoris$S_ADRD == 1),
  1,
  0
)

# Prioritize SveDem info, fallback to Dementia codes
o_amoris$Alzheimers <- ifelse(
  o_amoris$S_l_alz == 1 | o_amoris$S_e_alz == 1 | o_amoris$S_alz_vas == 1,
  1,
  ifelse(o_amoris$Dementia %in% alzheimers_codes, 1, 0)
)

o_amoris$Vascular <- ifelse(
  o_amoris$S_vas == 1,
  1,
  ifelse(o_amoris$Dementia %in% vascular_codes, 1, 0)
)

o_amoris$Other <- ifelse(
  o_amoris$S_oth == 1 | o_amoris$S_uns == 1 | o_amoris$S_par == 1 | o_amoris$S_fro == 1 | o_amoris$S_lew == 1,
  1,
  ifelse(o_amoris$Dementia %in% other_codes, 1, 0)
)

# Adds the people with no dementia diagnosis but ATC code to alzheimers, the people with dementia drugs have alzheimers
o_amoris$Alzheimers <- ifelse(o_amoris$ATC != "" & o_amoris$Dementia == "" & o_amoris$S_ADRD == "", 1, o_amoris$Alzheimers)
```

```{r}
# Income level as quartiles
# Step 1: Create quartile columns
o_amoris$ArbInk_quartile <- cut(
  o_amoris$ArbInk,
  breaks = quantile(o_amoris$ArbInk, probs = seq(0, 1, 0.25), na.rm = TRUE),
  include.lowest = TRUE,
  labels = 1:4
)

o_amoris$DispInk_quartile <- cut(
  o_amoris$DispInk,
  breaks = quantile(o_amoris$DispInk, probs = seq(0, 1, 0.25), na.rm = TRUE),
  include.lowest = TRUE,
  labels = 1:4
)
# Step 2: Create income_bins with NA if neither value exists
o_amoris$income_bins <- ifelse(
  !is.na(o_amoris$DispInk_quartile),  # If DispInk_quartile is available, use it
  as.numeric(as.character(o_amoris$DispInk_quartile)),  # Convert factor to numeric
  ifelse(
    !is.na(o_amoris$ArbInk_quartile),  # Otherwise, if ArbInk_quartile is available, use it
    as.numeric(as.character(o_amoris$ArbInk_quartile)),
    NA  # If neither is available, assign NA
  )
)



# Education level
o_amoris$NewUtbNiva <- ifelse(
  !is.na(o_amoris$Sun2000niva_old) & o_amoris$Sun2000niva_old != "*" & !is.na(as.numeric(o_amoris$Sun2000niva_old)), # Check if SUN2000_old is numeric and not '*'
  as.numeric(o_amoris$Sun2000niva_old),                                                                  # Use SUN2000_old if valid
  ifelse(
    !is.na(o_amoris$UtbNiva) & o_amoris$UtbNiva != "*" & !is.na(as.numeric(o_amoris$UtbNiva)),           # Check if UtbNiva is numeric and not '*'
    as.numeric(o_amoris$UtbNiva),                                                                    # Use UtbNiva as fallback
    NA                                                                                             # Otherwise, assign NA
  )
)

#missing <- sum(o_amoris$NewUtbNiva == "" | is.na(o_amoris$NewUtbNiva))
#unique(o_amoris$NewUtbNiva)



# Calculate age at FirstDate in years
o_amoris$AgeAtFirstDate <- as.numeric(difftime(o_amoris$firstDate, o_amoris$FODDATn, units = "days") / 365.25)
# Calculate age at LastDate in years
#o_amoris$AgeAtLastDate <- as.numeric(difftime(o_amoris$lastDate, o_amoris$FODDATn, units = "days") / 365.25)


### We want to check if lastDate is FDATUM for the people with nonempty ATC col (chr string) who are empty in Dementia (no chr string) and S_ADRD
# Start with the default: age at lastDate
o_amoris$AgeAtLastDate <- as.numeric(difftime(o_amoris$lastDate, o_amoris$FODDATn, units = "days")) / 365.25

# Identify the "meds-only" group (people missed by other sources)
meds_only <- o_amoris$ATC != "" & !is.na(o_amoris$ATC) &
  (o_amoris$Dementia == "" | is.na(o_amoris$Dementia)) &
  o_amoris$S_ADRD == 0

# For this group, recalculate age using FDATUM
o_amoris$AgeAtLastDate[meds_only] <- as.numeric(difftime(o_amoris$FDATUM[meds_only], o_amoris$FODDATn[meds_only], units = "days")) / 365.25

### Now condsider that we want whichever is first SveDem INDATUM or regular INDATUM
# Step 3: For everyone else – if either INDATUM or S_INDATUM exists, use the earliest one
earliest_dx <- pmin(o_amoris$INDATUM, o_amoris$S_INDATUM, na.rm = TRUE)

has_dx_date <- !is.na(earliest_dx)

o_amoris$AgeAtLastDate[has_dx_date] <- as.numeric(
  difftime(earliest_dx[has_dx_date], o_amoris$FODDATn[has_dx_date], units = "days")
) / 365.25
```

```{r}
#Decades (but first and last is 20 years)
# Extract decade from birth year
o_amoris$birthcohort <- paste0(substr(format(o_amoris$FODDATn, "%Y"), 1, 3), "0s")

# Manually adjust the first and last groups
o_amoris$birthcohort[o_amoris$birthcohort %in% c("1890s", "1900s","1910s")] <- "1890s-1910s"
o_amoris$birthcohort[o_amoris$birthcohort %in% c("1960s","1970s", "1980s", "1990s")] <- "1960s-1990s"

# Check the results
table(o_amoris$birthcohort)
```

## 1.4 Linking this to our longitudinal data
```{r}
# we want AgeAtFirstDate, AgeAtLastDate, income_bins, NewUtbNiva, ADRD, Alzheimers, Vascular, Other
#samplID
results_dt <- results_dt %>%
  left_join(o_amoris %>%
              select(AgeAtFirstDate, AgeAtLastDate, income_bins, NewUtbNiva, birthcohort,
                     ADRD, Alzheimers, Vascular, Other, sampleID),
            by = "sampleID")
```

## 1.5 Create the baseline for the longitudinal subset
```{r}
results_df <- results_dt %>%
  group_by(sampleID) %>%
  slice_min(order_by = age, n = 1, with_ties = FALSE) %>%
  ungroup()
```

## 1.6 Optional: input for descriptive tables 
```{r}
fortable1_RMA <- results_df
```

## 1.6 Beta function 
```{r}
calc_standardized_beta <- function(model, data, covariate) {
    beta <- coef(model)[covariate]
    sd_covariate <- sd(data[[covariate]], na.rm = TRUE)
    std_beta <- beta * sd_covariate
    return(std_beta)
}
```

# 2.0 Time to event analysis: association at baseline for the longitudinal subset

## 2.1 ADRD
```{r}
# BA measures to be tested
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Correct the initialization of HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),  # Repeat each BA measure twice
  outcome = rep("ADRD", times = length(BA_measures) * 2),
  model = rep(c("Sex-adjusted model", "Multivariable model"), times = length(BA_measures)),
  beta = NA, se = NA, HR = NA, CI_lower = NA, CI_upper = NA, p = NA, p_PH_BA = NA, global_p_PH_BA = NA,
  concordance = NA, N = NA, events = NA , std_beta = NA  # Add a column for the concordance index
)

for (j in 1:length(BA_measures)) {
  # Sex-adjusted model
  fit_gender <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, ADRD) ~", BA_measures[j], "+ gender")), data = results_df)
  coef_gender <- summary(fit_gender)
  row_index_gender <- (j * 2) - 1
  std_beta_gender <- calc_standardized_beta(fit_gender, results_df, BA_measures[j])
  HR_BA[row_index_gender, c(4, 5, 9, 12)] <- c(coef_gender$coefficients[1, c(1, 3, 5)], coef_gender$concordance[1])
  HR_BA[row_index_gender, c("HR", "CI_lower", "CI_upper")] <- coef_gender$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_gender, 10] <- cox.zph(fit_gender)$table[1, 3]
  HR_BA[row_index_gender, 11] <- cox.zph(fit_gender)$table["GLOBAL", "p"]
  HR_BA[row_index_gender, "N"] <- coef_gender$n
  HR_BA[row_index_gender, "std_beta"] <- std_beta_gender
  HR_BA[row_index_gender, "events"] <- fit_gender$nevent
  
  # Multivariable model
  fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, ADRD) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva ")), data = results_df) #birthcohort
  coef_multivariable <- summary(fit_multivariable)
  row_index_multivariable <- j * 2
  std_beta_multivariable <- calc_standardized_beta(fit_multivariable, results_df, BA_measures[j])
  HR_BA[row_index_multivariable, c(4, 5, 9, 12)] <- c(coef_multivariable$coefficients[1, c(1, 3, 5)], coef_multivariable$concordance[1])
  HR_BA[row_index_multivariable, c("HR", "CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_multivariable, 10] <- cox.zph(fit_multivariable)$table[1, 3]
  HR_BA[row_index_multivariable, 11] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
  HR_BA[row_index_multivariable, "N"] <- coef_multivariable$n
  HR_BA[row_index_multivariable, "std_beta"] <- std_beta_multivariable
  HR_BA[row_index_multivariable, "events"] <- fit_multivariable$nevent
}

ADRD_results <- HR_BA

# Print the modified HR_BA_adrd data frame
print(HR_BA)
```

## 2.2 Alzheimers
```{r}
# BA measures to be tested
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Correct the initialization of HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),  # Repeat each BA measure twice
  outcome = rep("Alzheimers", times = length(BA_measures) * 2),
  model = rep(c("Sex-adjusted model", "Multivariable model"), times = length(BA_measures)),
  beta = NA, se = NA, HR = NA, CI_lower = NA, CI_upper = NA, p = NA, p_PH_BA = NA, global_p_PH_BA = NA,
  concordance = NA, N = NA, events = NA , std_beta = NA  # Add a column for the concordance index
)

for (j in 1:length(BA_measures)) {
  # Sex-adjusted model
  fit_gender <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, Alzheimers) ~", BA_measures[j], "+ gender")), data = results_df)
  coef_gender <- summary(fit_gender)
  row_index_gender <- (j * 2) - 1
  std_beta_gender <- calc_standardized_beta(fit_gender, results_df, BA_measures[j])
  HR_BA[row_index_gender, c(4, 5, 9, 12)] <- c(coef_gender$coefficients[1, c(1, 3, 5)], coef_gender$concordance[1])
  HR_BA[row_index_gender, c("HR", "CI_lower", "CI_upper")] <- coef_gender$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_gender, 10] <- cox.zph(fit_gender)$table[1, 3]
  HR_BA[row_index_gender, 11] <- cox.zph(fit_gender)$table["GLOBAL", "p"]
  HR_BA[row_index_gender, "N"] <- coef_gender$n
  HR_BA[row_index_gender, "std_beta"] <- std_beta_gender
  HR_BA[row_index_gender, "events"] <- fit_gender$nevent
  
  # Multivariable model
  fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, Alzheimers) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva ")), data = results_df) #birthcohort
  coef_multivariable <- summary(fit_multivariable)
  row_index_multivariable <- j * 2
  std_beta_multivariable <- calc_standardized_beta(fit_multivariable, results_df, BA_measures[j])
  HR_BA[row_index_multivariable, c(4, 5, 9, 12)] <- c(coef_multivariable$coefficients[1, c(1, 3, 5)], coef_multivariable$concordance[1])
  HR_BA[row_index_multivariable, c("HR", "CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_multivariable, 10] <- cox.zph(fit_multivariable)$table[1, 3]
  HR_BA[row_index_multivariable, 11] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
  HR_BA[row_index_multivariable, "N"] <- coef_multivariable$n
  HR_BA[row_index_multivariable, "std_beta"] <- std_beta_multivariable
  HR_BA[row_index_multivariable, "events"] <- fit_multivariable$nevent
}

Alzh_results <- HR_BA

# Print the modified HR_BA_adrd data frame
print(HR_BA)
```

## 2.3 Vascular
```{r}
# BA measures to be tested
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Correct the initialization of HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),  # Repeat each BA measure twice
  outcome = rep("Vascular", times = length(BA_measures) * 2),
  model = rep(c("Sex-adjusted model", "Multivariable model"), times = length(BA_measures)),
  beta = NA, se = NA, HR = NA, CI_lower = NA, CI_upper = NA, p = NA, p_PH_BA = NA, global_p_PH_BA = NA,
  concordance = NA, N = NA, events = NA , std_beta = NA  # Add a column for the concordance index
)

for (j in 1:length(BA_measures)) {
  # Sex-adjusted model
  fit_gender <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, Vascular) ~", BA_measures[j], "+ gender")), data = results_df)
  coef_gender <- summary(fit_gender)
  row_index_gender <- (j * 2) - 1
  std_beta_gender <- calc_standardized_beta(fit_gender, results_df, BA_measures[j])
  HR_BA[row_index_gender, c(4, 5, 9, 12)] <- c(coef_gender$coefficients[1, c(1, 3, 5)], coef_gender$concordance[1])
  HR_BA[row_index_gender, c("HR", "CI_lower", "CI_upper")] <- coef_gender$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_gender, 10] <- cox.zph(fit_gender)$table[1, 3]
  HR_BA[row_index_gender, 11] <- cox.zph(fit_gender)$table["GLOBAL", "p"]
  HR_BA[row_index_gender, "N"] <- coef_gender$n
  HR_BA[row_index_gender, "std_beta"] <- std_beta_gender
  HR_BA[row_index_gender, "events"] <- fit_gender$nevent
  
  # Multivariable model
  fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, Vascular) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva")), data = results_df)
  coef_multivariable <- summary(fit_multivariable)
  row_index_multivariable <- j * 2
  std_beta_multivariable <- calc_standardized_beta(fit_multivariable, results_df, BA_measures[j])
  HR_BA[row_index_multivariable, c(4, 5, 9, 12)] <- c(coef_multivariable$coefficients[1, c(1, 3, 5)], coef_multivariable$concordance[1])
  HR_BA[row_index_multivariable, c("HR", "CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_multivariable, 10] <- cox.zph(fit_multivariable)$table[1, 3]
  HR_BA[row_index_multivariable, 11] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
  HR_BA[row_index_multivariable, "N"] <- coef_multivariable$n
  HR_BA[row_index_multivariable, "std_beta"] <- std_beta_multivariable
  HR_BA[row_index_multivariable, "events"] <- fit_multivariable$nevent
}

Vasc_results <- HR_BA

# Print the modified HR_BA_adrd data frame
print(HR_BA)
```

## 2.4 Other
```{r}
# BA measures to be tested
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Correct the initialization of HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),  # Repeat each BA measure twice
  outcome = rep("Other", times = length(BA_measures) * 2),
  model = rep(c("Sex-adjusted model", "Multivariable model"), times = length(BA_measures)),
  beta = NA, se = NA, HR = NA, CI_lower = NA, CI_upper = NA, p = NA, p_PH_BA = NA, global_p_PH_BA = NA,
  concordance = NA, N = NA, events = NA , std_beta = NA  # Add a column for the concordance index
)

for (j in 1:length(BA_measures)) {
  # Sex-adjusted model
  fit_gender <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, Other) ~", BA_measures[j], "+ gender")), data = results_df)
  coef_gender <- summary(fit_gender)
  row_index_gender <- (j * 2) - 1
  std_beta_gender <- calc_standardized_beta(fit_gender, results_df, BA_measures[j])
  HR_BA[row_index_gender, c(4, 5, 9, 12)] <- c(coef_gender$coefficients[1, c(1, 3, 5)], coef_gender$concordance[1])
  HR_BA[row_index_gender, c("HR", "CI_lower", "CI_upper")] <- coef_gender$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_gender, 10] <- cox.zph(fit_gender)$table[1, 3]
  HR_BA[row_index_gender, 11] <- cox.zph(fit_gender)$table["GLOBAL", "p"]
  HR_BA[row_index_gender, "N"] <- coef_gender$n
  HR_BA[row_index_gender, "std_beta"] <- std_beta_gender
  HR_BA[row_index_gender, "events"] <- fit_gender$nevent
  
  # Multivariable model
  fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, Other) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva ")), data = results_df)
  coef_multivariable <- summary(fit_multivariable)
  row_index_multivariable <- j * 2
  std_beta_multivariable <- calc_standardized_beta(fit_multivariable, results_df, BA_measures[j])
  HR_BA[row_index_multivariable, c(4, 5, 9, 12)] <- c(coef_multivariable$coefficients[1, c(1, 3, 5)], coef_multivariable$concordance[1])
  HR_BA[row_index_multivariable, c("HR", "CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_multivariable, 10] <- cox.zph(fit_multivariable)$table[1, 3]
  HR_BA[row_index_multivariable, 11] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
  HR_BA[row_index_multivariable, "N"] <- coef_multivariable$n
  HR_BA[row_index_multivariable, "std_beta"] <- std_beta_multivariable
  HR_BA[row_index_multivariable, "events"] <- fit_multivariable$nevent
}

Othe_results <- HR_BA

# Print the modified HR_BA_adrd data frame
print(HR_BA)
```

## 2.5 Binding and saving
```{r}
# Combine all data frames into one large data frame
combined_results <- rbind(
  Alzh_results,
  ADRD_results,
  Othe_results,
  Vasc_results #for multi birth cohort was removed 
)

```

```{r}
write.table(as.data.frame(combined_results), "combined_results_ADRD_RMA.csv", sep = ";",
            dec = ",",
            row.names = FALSE,
            quote = FALSE,
            fileEncoding = "Latin1")
```

# 3.0 Rate of change or intercept?

## 3.1 Making slopes (PhenoAge)
```{r}
library(data.table)

dt <- as.data.table(results_dt)

# Compute per-individual slope and intercept with age centered at first measurement
slopes_dt <- dt[, if (.N >= 3) {
  age0 <- min(age, na.rm = TRUE)  # baseline CA
  model <- lm(phenoage_advance_sd ~ I(age - age0))
  .(slope = coef(model)[2], intercept = coef(model)[1])
}, by = sampleID]

# Merge back slope/intercept to all rows for eligible sampleIDs
results_phe <- merge(dt, slopes_dt, by = "sampleID", all.x = FALSE, all.y = TRUE)
```

## 3.2 Time to event modeling 
```{r}
slope_phe <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, ADRD) ~", "slope + intercept", "+ gender")), data = results_phe)
  slope_phe_sum <- summary(slope_phe)
```

## 3.3 Making slopes (KDM)
```{r}
library(data.table)

dt <- as.data.table(results_dt)

# Compute per-individual slope and intercept with age centered at first measurement
slopes_dt <- dt[, if (.N >= 3) {
  age0 <- min(age, na.rm = TRUE)  # baseline CA
  model <- lm(kdm_advance_sd ~ I(age - age0))
  .(slope = coef(model)[2], intercept = coef(model)[1])
}, by = sampleID]

# Merge back slope/intercept to all rows for eligible sampleIDs
results_kdm <- merge(dt, slopes_dt, by = "sampleID", all.x = FALSE, all.y = TRUE)
```

## 3.4 Time to event modeling 
```{r}
slope_kdm <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, ADRD) ~", "slope + intercept", "+ gender")), data = results_kdm)
  slope_kdm_sum <- summary(slope_kdm)
```
