---
title: "Calculating_BA_L"
output: html_notebook
author: Karolina Gustavsson
date: 2026-04-10
---

Input data: AMORIS (longitudinal) and sampleIDs from the baseline training set. An individual can have several rows if they have repeated measurements. 

Important note: We want the training set to be the same as for AMORIS (baseline).

# 1.0 Preparing variables

## 1.1 Loading packages
```{r}
options(repos = "http://nexus.ki.se/repository/cran.r-project.org")
.libPaths("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/shun_trial/lib")
source("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/shun_trial/function.R")

library(haven)
library(broom)
library(dplyr)
library(flexsurv)
library(ggplot2)
library(htmlTable)
library(survival)
library(survey)
library(tidyr)
library(rmarkdown)
```

## 1.2 Load data 
```{r}
n_amoris <- read_sas("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/Data_ADRD/AMORIS_20240417.sas7bdat") 
```

```{r}
# Define the path, IDs for individuals selected at baseline
path_to_file <- "W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/PhD_studies/Study_1/Script/Current_version/Baseline/sampleIDs_baseline.csv"

# Read the file with custom settings
sampleID_df <- read.table(path_to_file,
                          sep = ";",
                          dec = ",",
                          header = TRUE,
                          fileEncoding = "Latin1")
```

## 1.3 Defining our selected biomarkers
```{r}
biomarkers = c("S_Alb", "S_Krea", "TC", "TG", "S_K", "S_Urea", "S_LD", "S_Ca", "S_Urat", "S_FAMN", "S_Hapt", "fS_Gluk") 
n_amoris <- n_amoris[complete.cases(n_amoris[, biomarkers]), ]
```

## 1.4 Variables required by BioAge
```{r}
colnames(n_amoris)[colnames(n_amoris) == "Kon"] <- "gender"
colnames(n_amoris)[colnames(n_amoris) == "Alder"] <- "age" #already called age
colnames(n_amoris)[colnames(n_amoris) == "Id"] <- "sampleID"

n_amoris <- n_amoris %>%
  mutate(status = ifelse(is.na(DODSDATn), 0, 1)) %>%
  mutate(time = as.numeric(difftime(lastDate, firstDate, units = "weeks")) / 4.33)
```

## 1.5 Selecting the sampleIDs at baseline for our longituidinal samples
```{r}
sampleID_df <- sampleID_df %>%
  rename(sampleID = to1po2.sampleID)

N_AMORIS <- n_amoris %>%
  semi_join(sampleID_df, by = "sampleID")
```

## 1.6 Age resolution
```{r}
#save integer age col before we relace it
N_AMORIS$age_int <- N_AMORIS$age

# Age at last test for biomarkers/BA
N_AMORIS <- N_AMORIS %>%
  mutate(age = as.numeric(difftime(Provtagningsdatum, FODDATn, units = "weeks")) / 52.25)
  
N_AMORIS <- N_AMORIS %>%
  mutate(age = round(age, 2))
```

## 1.7 Loading test set
```{r}
# Define the path
path_to_file <- "W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/PhD_studies/Study_1/Script/Current_version/Baseline/training_set_baseline.csv"

# Read the file with custom settings
train_baseline <- read.table(path_to_file,
                          sep = ";",
                          dec = ",",
                          header = TRUE,
                          fileEncoding = "Latin1")
```

## 1.8 Separate test set from training set
```{r}
library(dplyr)

test_longitud <- N_AMORIS %>%
  anti_join(train_baseline, by = "sampleID")
```

## 1.9 Visual inspection
```{r}
# Load the ggplot2 package
library(ggplot2)

# Assuming 'amoris' DataFrame is already loaded and 'Kon' is your gender column with 1 for male and 2 for female
# Convert the numeric codes to a factor to improve plot readability
test_longitud$Sex <- factor(test_longitud$gender, levels = c(1, 2), labels = c("Men", "Women"))

# Create overlapping histograms with transparency
a_test_set <- ggplot(test_longitud, aes(x = age_int, fill = Sex)) +
  geom_histogram(position = "identity", alpha = 0.5, binwidth = 1) +  # binwidth can be adjusted based on data
  scale_fill_manual(values = c("#6BAED6", "#E78AC3")) +  # Colors can be adjusted
  labs(x = "Age", y = "Count") +
  theme_minimal()

a_test_set
```

```{r}
png("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/PhD_studies/Study_1/Script/Current_version/Longitudinal/a_test_set_l_bluepink.png", 
    width = 6, height = 4, units = 'in', res = 600)
print(a_test_set)  # Send plot to file
dev.off()  # Close device
```

# 2.0 BioAge (KDM, PhenoAge)

## 2.1 KDM
```{r}
## KDM
train_f = kdm_calc(train_baseline %>%
                     filter(gender ==2),
                    #filter(age >=  25 & age <= 75, gender ==2),
                   biomarkers, fit = NULL, s_ba2 = NULL)
train_m = kdm_calc(train_baseline %>%
                     filter(gender ==1),
                    #filter(age >= 25 & age <= 75, gender ==1),
                   biomarkers, fit = NULL, s_ba2 = NULL)
kdm_f = kdm_calc(test_longitud %>%
                   filter (gender ==2), #age >= 25 & age <= 75
                 biomarkers,
                 fit = train_f$fit,
                 s_ba2 = train_f$fit$s_ba2)
kdm_m = kdm_calc(test_longitud %>%
                   filter (gender ==1), #age >= 25 & age <= 75
                 biomarkers,
                 fit = train_m$fit,
                 s_ba2 = train_m$fit$s_ba2)
to1po2_kdm = rbind(kdm_f$data,kdm_m$data)

rm(train_f)
rm(train_m)
rm(kdm_f)
rm(kdm_m)
```

## 2.2 Phenoage
```{r}
train = phenoage_calc(train_baseline, biomarkers)
#Project into the AMORIS_2
phenoage = phenoage_calc(test_longitud, biomarkers, fit = train$fit)
#Extract phenoage dataset, trained on 1 projected on 2 = to1po2, previously data
to1po2_phe = (phenoage$data)

rm(phenoage)
rm(train)
```

## 2.3 Merging
```{r}
# since it's longitudinal we need to consider how we merge the data 
to1po2 <- left_join(test_longitud, to1po2_kdm[, c("sampleID", "Provtagningsdatum", "kdm", "kdm_advance")], 
                    by = c("sampleID", "Provtagningsdatum")) %>%
  left_join(., to1po2_phe[, c("sampleID", "Provtagningsdatum", "phenoage", "phenoage_advance")], 
            by = c("sampleID", "Provtagningsdatum"))
```

## 2.4 Visual inspection of KDM and PhenoAge
```{r}
 agevar_ap = c("kdm","phenoage")
 #prepare labels
 label_ap = c("KDM AMORIS",
              "PhenoAge AMORIS")
 #plot age vs bioage
 c_modified_kdm_phe <- plot_ba(to1po2, agevar_ap, label_ap)
 
 print(c_modified_kdm_phe)
```

```{r}
png("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/PhD_studies/Study_1/Script/Current_version/Longitudinal/c_corr_kdm_phe_l.png", 
    width = 8, height = 4, units = 'in', res = 600)
print(c_modified_kdm_phe)  # Send plot to file
dev.off()  # Close device
```

## 2.6 Correlation with BA
```{r}
get_axis_type <- function(labels){
  return(rep("float", length(labels)) %>% setNames(names(labels))) # Create function to generate axis_type variables for BAA plots
}

png("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/PhD_studies/Study_1/Script/Current_version/Longitudinal/c_corrplot_l_good.png", 
    width = 6, height = 4.5, units = 'in', res = 600)

plot_baa(data=to1po2,
         agevar=names(c("kdm"="KDMAge", "kdm_advance"="KDM\nAge\nGap", "phenoage"="PhenoAge","phenoage_advance"="PhenoAge\nAge\nGap", "age"="Chronological\nage")), 
         label = c("kdm" = "KDMAge\n","kdm_advance"="KDM\nAge\nGap",
                        "phenoage" = "PhenoAge", "phenoage_advance"="PhenoAge\nAge\nGap",
                        "age"="Chronological\nage"),
         axis_type= c("kdm" = "float",
                            "kdm_advance" = "float",
                            "phenoage" = "float",
                            "phenoage_advance" = "float",
                            "age" = "float"))

```
