---
title: "Trajectories_L"
output: html_notebook
author: Karolina Gustavsson
date: 2026-05-14
---

Input data: AMORIS (longitudinal) dvs results_dt från BA_rate_L.rmd

# 1.0 Preparing

## 1.1 Selecting what to plot
```{r}
to1po2_women <- results_dt[results_dt$gender == 2, ]
to1po2_men <- results_dt[results_dt$gender == 1, ]
```

## Optional: health checkups
```{r}
results_FH_dt <- results_dt[results_dt$debiteringskod %in% c("F", "H"), ] #Health checkups

to1po2_FH_women <- results_dt[results_dt$debiteringskod %in% c("F", "H") & results_dt$gender == 2, ]

to1po2_FH_men <- results_dt[results_dt$debiteringskod %in% c("F", "H") & results_dt$gender == 1, ]
```

# 2.0 Calculating trajectories binned by 5 years

## 2.1 Phenoage
```{r}
# Load required packages
library(dplyr)
library(sandwich)
library(lmtest)

# --- Step 1: Prepare the data ---
to1po2_clean <- to1po2_women %>% #results_dt for all or to1po2_FH for only those with health checkups
  filter(!is.na(age), !is.na(phenoage)) %>%
  group_by(sampleID) %>%
  filter(n() >= 3) %>%
  ungroup() %>%
  mutate(
    age_bin = cut(
      age,
      #breaks = seq(floor(min(age, na.rm = TRUE)),
      #             ceiling(max(age, na.rm = TRUE)),
      #             by = 5),
      breaks = seq(17, ceiling(max(age, na.rm = TRUE)), by = 5),
      include.lowest = TRUE,
      right = FALSE
    )
  )

# --- Step 2: Fit the model with interaction terms ---
fit <- lm(phenoage ~ age + age:age_bin, data = to1po2_clean)
coefs <- coef(fit)
vcov_clustered <- vcovCL(fit, cluster = to1po2_clean$sampleID)

# --- Step 3: Global Wald test ---
interaction_terms <- grep("age:age_bin", names(coefs))
estimates <- coefs[interaction_terms]
V <- vcov_clustered[interaction_terms, interaction_terms]

wald_stat <- t(estimates) %*% solve(V) %*% estimates
df <- length(estimates)
p_value <- 1 - pchisq(wald_stat, df)

# --- Step 4: Compute slope estimates and confidence intervals ---
z <- 1.96
slope_table <- data.frame(
  age_bin = levels(to1po2_clean$age_bin),
  slope = NA,
  se = NA,
  lower_CI = NA,
  upper_CI = NA
)
library(stringr)

# Extract start and stop values from age_bin labels
bounds <- str_extract_all(slope_table$age_bin, "[0-9.]+")
slope_table$start <- as.numeric(sapply(bounds, `[`, 1))
slope_table$stop <- as.numeric(sapply(bounds, `[`, 2))

# Reference bin (first bin)
base_slope <- coefs["age"]
base_se <- sqrt(vcov_clustered["age", "age"])

slope_table[1, "slope"] <- base_slope
slope_table[1, "se"] <- base_se
slope_table[1, "lower_CI"] <- base_slope - z * base_se
slope_table[1, "upper_CI"] <- base_slope + z * base_se

# All other bins
for (i in 2:nrow(slope_table)) {
  bin <- slope_table$age_bin[i]
  term <- paste0("age:age_bin", bin)

  est <- base_slope + coefs[term]

  var_base <- vcov_clustered["age", "age"]
  var_int  <- vcov_clustered[term, term]
  covar    <- 2 * vcov_clustered["age", term]
  se <- sqrt(var_base + var_int + covar)

  slope_table[i, "slope"] <- est
  slope_table[i, "se"] <- se
  slope_table[i, "lower_CI"] <- est - z * se
  slope_table[i, "upper_CI"] <- est + z * se
}

# --- Output ---
cat("===== GLOBAL TEST (Wald) =====\n")
cat("Wald statistic:", round(wald_stat, 3), "\n")
cat("Degrees of freedom:", df, "\n")
cat("P-value:", paste0("< ", .Machine$double.eps), "\n\n")

cat("===== SLOPE ESTIMATES BY AGE BIN =====\n")
print(slope_table)

slope_table_all_phe <- slope_table
```

```{r}
write.table(as.data.frame(slope_table),
            file = "trajectory_table_phe.csv",
            sep = ";",
            dec = ",",
            row.names = FALSE,
            quote = FALSE,
            fileEncoding = "Latin1") 
```

```{r}
# Create a one-row summary table with global Wald test output
wald_summary <- data.frame(
  wald_statistic = as.numeric(wald_stat),
  degrees_of_freedom = df,
  p_value = as.numeric(p_value)
)

print(wald_summary)
```

```{r}
write.table(as.data.frame(wald_summary),
            file = "trajectory_table_wald_phe_FH.csv",
            sep = ";",
            dec = ",",
            row.names = FALSE,
            quote = FALSE,
            fileEncoding = "Latin1") 
```

## 2.2 Creating figure (PhenoAge)
```{r}
library(dplyr)
library(ggplot2)
library(stringr)

# Assume slope_table and clean_df already exist
# Extract global intercept from model
intercept <- coef(fit)["(Intercept)"]

# Construct CI-based rotated rectangles
plot_data <- slope_table %>%
  rowwise() %>%
  mutate(
    x0 = start,
    x1 = stop,
    y0 = intercept + slope * x0,
    y1 = intercept + slope * x1,

    dx = x1 - x0,
    dy = y1 - y0,
    L = sqrt(dx^2 + dy^2),

    nx = -dy / L,
    ny = dx / L,

    slope_ci = (upper_CI - lower_CI) / 2,
    w = slope_ci * dx / sqrt(1 + slope^2),

    xA = x0 + w * nx,
    yA = y0 + w * ny,
    xB = x1 + w * nx,
    yB = y1 + w * ny,
    xC = x1 - w * nx,
    yC = y1 - w * ny,
    xD = x0 - w * nx,
    yD = y0 - w * ny,

    label_x = (x0 + x1) / 2,
    label_y = (y0 + y1) / 2,
    #label = round(slope, 2),
    label = sprintf("%.2f", slope),
    group = paste0("bin_", start)
  ) %>%
  ungroup()

plot_data <- plot_data %>%
  filter(stop <= 82)

# Long-format polygon
polygon_df <- plot_data %>%
  rowwise() %>%
  do(data.frame(
    group = .$group,
    x = c(.$xA, .$xB, .$xC, .$xD),
    y = c(.$yA, .$yB, .$yC, .$yD)
  )) %>%
  ungroup()

# Optional spaghetti overlay toggle
show_spaghetti <- TRUE

#plot_data$label <- sprintf("%.2f", plot_data$label)

# Start ggplot
phenoage_plot <- ggplot()

# Add spaghetti first (behind everything)
if (show_spaghetti) {
  set.seed(42)
  subset_ids <- sample(unique(to1po2_clean$sampleID), floor(0.01 * length(unique(to1po2_clean$sampleID))))
  spaghetti_df <- to1po2_clean %>% filter(sampleID %in% subset_ids)

  # phenoage_plot <- phenoage_plot +
  #   geom_line(
  #     data = spaghetti_df,
  #     aes(x = age, y = phenoage, group = sampleID),
  #     color = "#0072B2", alpha = 0.1, linewidth = 0.3
  #   ) +
  #   geom_point(
  #     data = spaghetti_df,
  #     aes(x = age, y = phenoage),
  #     color = "#0072B2", alpha = 0.1, size = 0.5
  #   )
  
  phenoage_plot <- phenoage_plot +
  ## lines per participant
  geom_line(
    data = spaghetti_df,
    aes(x = age, y = phenoage, group = sampleID),
    color = "#0072B2", alpha = 0.30, linewidth = 0.3, show.legend = FALSE
  ) +
  ## tiny observation dots
  geom_point(
    data = spaghetti_df,
    aes(x = age, y = phenoage),
    color = "#0072B2", alpha = 0.50, size = 0.10, show.legend = FALSE
  )
}

# Add rotated CI rectangles, lines, and slope labels
phenoage_plot <- phenoage_plot +
  geom_polygon(
    data = polygon_df,
    aes(x = x, y = y, group = group),
    fill = "darkred", alpha = 0.3
  ) +
  geom_segment(
    data = plot_data,
    aes(x = x0, xend = x1, y = y0, yend = y1),
    color = "darkred", size = 1
  ) +
  # geom_text(
  #   data = plot_data,
  #   aes(x = label_x, y = label_y, label = label),
  #   size = 3,
  #   fontface = "bold",
  #   color = "black",
  #   vjust = -2,   # Raise label slightly above the segment
  #   hjust = 0.6   # Center horizontally
  # )
  ggtext::geom_richtext(
  data = plot_data,
  aes(x = label_x, y = label_y, label = label),
  size = 3,
  fontface = "bold",
  color = "black",
  fill = alpha("white",0.4),       # background
  label.color = NA,     # remove border around label
  vjust = -2,
  hjust = 0.6
) +

#phenoage_plot <- phenoage_plot +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  coord_fixed(ratio = 1, xlim = c(15, 85), ylim = c(15, 100)) +
  scale_x_continuous(breaks = seq(10, 85, by = 10)) +
  scale_y_continuous(breaks = seq(10, 100, by = 10)) +
  labs(x = "Chronological Age", y = "PhenoAge") +
  theme_minimal()

print(phenoage_plot)
```

```{r}
png("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/PhD_studies/Study_1/Script/Current_version/Longitudinal/trajectory_all_phe_v4_spag.png", 
    width = 6, height = 6, units = 'in', res = 1000)
print(phenoage_plot)  # Send plot to file
dev.off()  # Close device
```

## 2.3 KDM
```{r}
# Load required packages
library(dplyr)
library(sandwich)
library(lmtest)

# --- Step 1: Prepare the data ---
to1po2_clean <- results_dt %>% #results_dt for all or to1po2_FH for those with health checkups, MAIN results_dt
  filter(!is.na(age), !is.na(kdm)) %>%
  group_by(sampleID) %>%
  filter(n() >= 3) %>%
  ungroup() %>%
  mutate(
    age_bin = cut(
      age,
      #breaks = seq(floor(min(age, na.rm = TRUE)),
      #             ceiling(max(age, na.rm = TRUE)),
      #             by = 5),
      breaks = seq(17, ceiling(max(age, na.rm = TRUE)), by = 5),
      include.lowest = TRUE,
      right = FALSE
    )
  )

# --- Step 2: Fit the model with interaction terms ---
fit <- lm(kdm ~ age + age:age_bin, data = to1po2_clean)
coefs <- coef(fit)
vcov_clustered <- vcovCL(fit, cluster = to1po2_clean$sampleID)

# --- Step 3: Global Wald test ---
interaction_terms <- grep("age:age_bin", names(coefs))
estimates <- coefs[interaction_terms]
V <- vcov_clustered[interaction_terms, interaction_terms]

wald_stat <- t(estimates) %*% solve(V) %*% estimates
df <- length(estimates)
p_value <- 1 - pchisq(wald_stat, df)

# --- Step 4: Compute slope estimates and confidence intervals ---
z <- 1.96
slope_table <- data.frame(
  age_bin = levels(to1po2_clean$age_bin),
  slope = NA,
  se = NA,
  lower_CI = NA,
  upper_CI = NA
)

library(stringr)

# Extract start and stop values from age_bin labels
bounds <- str_extract_all(slope_table$age_bin, "[0-9.]+")
slope_table$start <- as.numeric(sapply(bounds, `[`, 1))
slope_table$stop <- as.numeric(sapply(bounds, `[`, 2))

# Reference bin (first bin)
base_slope <- coefs["age"]
base_se <- sqrt(vcov_clustered["age", "age"])

slope_table[1, "slope"] <- base_slope
slope_table[1, "se"] <- base_se
slope_table[1, "lower_CI"] <- base_slope - z * base_se
slope_table[1, "upper_CI"] <- base_slope + z * base_se

# All other bins
for (i in 2:nrow(slope_table)) {
  bin <- slope_table$age_bin[i]
  term <- paste0("age:age_bin", bin)
  
  est <- base_slope + coefs[term]
  
  var_base <- vcov_clustered["age", "age"]
  var_int  <- vcov_clustered[term, term]
  covar    <- 2 * vcov_clustered["age", term]
  se <- sqrt(var_base + var_int + covar)
  
  slope_table[i, "slope"] <- est
  slope_table[i, "se"] <- se
  slope_table[i, "lower_CI"] <- est - z * se
  slope_table[i, "upper_CI"] <- est + z * se
}

# --- Step 5: Ensure numeric and round all numeric columns ---
slope_table <- slope_table %>%
  mutate(across(where(is.numeric), ~ round(., 4)))

# --- Output ---
cat("===== GLOBAL TEST (Wald) =====\n")
cat("Wald statistic:", round(wald_stat, 3), "\n")
cat("Degrees of freedom:", df, "\n")
cat("P-value:", paste0("< ", .Machine$double.eps), "\n\n")

cat("===== SLOPE ESTIMATES BY AGE BIN =====\n")
print(slope_table)
slope_table_all_kdm <- slope_table
```

```{r}
write.table(as.data.frame(slope_table_all_kdm),
            file = "trajectory_table_kdm_FH.csv",
            sep = ";",
            dec = ",",
            row.names = FALSE,
            quote = FALSE,
            fileEncoding = "Latin1") 
```

```{r}
# Create a one-row summary table with global Wald test output
wald_summary <- data.frame(
  wald_statistic = as.numeric(wald_stat),
  degrees_of_freedom = df,
  p_value = as.numeric(p_value)
)

print(wald_summary)
```

```{r}
write.table(as.data.frame(wald_summary),
            file = "trajectory_table_wald_kdm_FH.csv",
            sep = ";",
            dec = ",",
            row.names = FALSE,
            quote = FALSE,
            fileEncoding = "Latin1") 
```

## 2.4 Creating figure (KDM)
```{r}
library(dplyr)
library(ggplot2)
library(stringr)

# Assume slope_table and to1po2_clean already exist
# Extract global intercept from model
intercept <- coef(fit)["(Intercept)"]

# Construct CI-based rotated rectangles
plot_data <- slope_table %>%
  rowwise() %>%
  mutate(
    x0 = start,
    x1 = stop,
    y0 = intercept + slope * x0,
    y1 = intercept + slope * x1,

    dx = x1 - x0,
    dy = y1 - y0,
    L = sqrt(dx^2 + dy^2),

    nx = -dy / L,
    ny =  dx / L,

    slope_ci = (upper_CI - lower_CI) / 2,
    w = slope_ci * dx / sqrt(1 + slope^2),

    xA = x0 + w * nx,
    yA = y0 + w * ny,
    xB = x1 + w * nx,
    yB = y1 + w * ny,
    xC = x1 - w * nx,
    yC = y1 - w * ny,
    xD = x0 - w * nx,
    yD = y0 - w * ny,

    label_x = (x0 + x1) / 2,
    label_y = (y0 + y1) / 2,
    label = round(slope, 2),
    group = paste0("bin_", start)
  ) %>%
  ungroup()

plot_data <- plot_data %>%
  filter(stop <= 82)

# Long-format polygon
polygon_df <- plot_data %>%
  rowwise() %>%
  do(data.frame(
    group = .$group,
    x = c(.$xA, .$xB, .$xC, .$xD),
    y = c(.$yA, .$yB, .$yC, .$yD)
  )) %>%
  ungroup()

# Optional spaghetti overlay toggle
show_spaghetti <- TRUE

plot_data$label <- sprintf("%.2f", plot_data$label)

# Start ggplot
kdm_plot <- ggplot()

# Add spaghetti first (behind everything)
if (show_spaghetti) {
  set.seed(42)
  subset_ids <- sample(unique(to1po2_clean$sampleID), floor(0.01 * length(unique(to1po2_clean$sampleID))))
  spaghetti_df <- to1po2_clean %>% filter(sampleID %in% subset_ids)

#   kdm_plot <- kdm_plot +
#     geom_line(
#       data = spaghetti_df,
#       aes(x = age, y = kdm, group = sampleID),
#       color = "#0072B2", alpha = 0.1, linewidth = 0.3
#     ) +
#     geom_point(
#       data = spaghetti_df,
#       aes(x = age, y = kdm),
#       color = "#0072B2", alpha = 0.1, size = 0.5
#     )
  
    kdm_plot <- kdm_plot +
  ## lines per participant
  geom_line(
    data = spaghetti_df,
    aes(x = age, y = kdm, group = sampleID),
    color = "#0072B2", alpha = 0.30, linewidth = 0.3, show.legend = FALSE
  ) +
  ## tiny observation dots
  geom_point(
    data = spaghetti_df,
    aes(x = age, y = kdm),
    color = "#0072B2", alpha = 0.50, size = 0.10, show.legend = FALSE
  )
 }

# Add rotated CI rectangles, lines, and slope labels
kdm_plot <- kdm_plot +
  geom_polygon(
    data = polygon_df,
    aes(x = x, y = y, group = group),
    fill = "darkred", alpha = 0.3
  ) +
  geom_segment(
    data = plot_data,
    aes(x = x0, xend = x1, y = y0, yend = y1),
    color = "darkred", size = 1
  ) +
  # geom_text(
  #   data = plot_data,
  #   aes(x = label_x, y = label_y, label = label),
  #   color = "black", size = 3
  # ) +
  # geom_text(
  #   data = plot_data,
  #   aes(x = label_x, y = label_y, label = label),
  #   size = 3,
  #   fontface = "bold",
  #   color = "black",
  #   vjust = -2,  # Raise label slightly above the segment
  #   hjust = 0.6    # Center horizontally
  # ) +
ggtext::geom_richtext(
  data = plot_data,
  aes(x = label_x, y = label_y, label = label),
  size = 3,
  fontface = "bold",
  color = "black",
  fill = alpha("white",0.4),       # background
  label.color = NA,     # remove border around label
  vjust = -2,
  hjust = 0.6
) +
  
  
  
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  coord_fixed(ratio = 1, xlim = c(15, 85), ylim = c(15, 100)) +
  scale_x_continuous(breaks = seq(10, 80, by = 10)) +
  scale_y_continuous(breaks = seq(10, 100, by = 10)) +
  labs(x = "Chronological Age", y = "KDM") +
  theme_minimal()

print(kdm_plot)

```

```{r}
png("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/PhD_studies/Study_1/Script/Current_version/Longitudinal/trajectory_kdm_men_v4_spag.png", 
    width = 6, height = 6, units = 'in', res = 1000)
print(kdm_plot)  # Send plot to file
dev.off()  # Close device
```

## 2.5 Creating figure colored by density (PhenoAge)
```{r}
library(dplyr)
library(ggplot2)
library(stringr)
library(MASS)
library(fields)

# Assume slope_table, to1po2_clean, and fit already exist

# Extract global intercept from model
intercept <- coef(fit)["(Intercept)"]

# Construct CI-based rotated rectangles
plot_data <- slope_table %>%
  rowwise() %>%
  mutate(
    x0 = start,
    x1 = stop,
    y0 = intercept + slope * x0,
    y1 = intercept + slope * x1,

    dx = x1 - x0,
    dy = y1 - y0,
    L = sqrt(dx^2 + dy^2),

    nx = -dy / L,
    ny = dx / L,

    slope_ci = (upper_CI - lower_CI) / 2,
    w = slope_ci * dx / sqrt(1 + slope^2),

    xA = x0 + w * nx,
    yA = y0 + w * ny,
    xB = x1 + w * nx,
    yB = y1 + w * ny,
    xC = x1 - w * nx,
    yC = y1 - w * ny,
    xD = x0 - w * nx,
    yD = y0 - w * ny,

    label_x = (x0 + x1) / 2,
    label_y = (y0 + y1) / 2,
    label = sprintf("%.2f", slope),
    group = paste0("bin_", start)
  ) %>%
  ungroup()

plot_data <- plot_data %>%
  filter(stop <= 82)

# Long-format polygon
polygon_df <- plot_data %>%
  rowwise() %>%
  do(data.frame(
    group = .$group,
    x = c(.$xA, .$xB, .$xC, .$xD),
    y = c(.$yA, .$yB, .$yC, .$yD)
  )) %>%
  ungroup()

# Optional spaghetti overlay toggle
show_spaghetti <- TRUE

# Start ggplot
phenoage_plot <- ggplot()

# Add spaghetti first (behind everything)
if (show_spaghetti) {
  set.seed(42)
  subset_ids <- sample(
    unique(to1po2_clean$sampleID),
    floor(0.01 * length(unique(to1po2_clean$sampleID)))
  )

  spaghetti_df <- to1po2_clean %>%
    filter(sampleID %in% subset_ids)

  # Same density approach as your previous KDM plot, adapted to phenoage
  add_kde_density <- function(df, x = "age", y = "phenoage", n = 250) {
    xx <- df[[x]]
    yy <- df[[y]]

    ok <- is.finite(xx) & is.finite(yy)

    out <- df
    out$dens <- NA_real_

    if (sum(ok) < 2 || length(unique(xx[ok])) < 2 || length(unique(yy[ok])) < 2) {
      return(out)
    }

    kde <- MASS::kde2d(xx[ok], yy[ok], n = n)
    surf <- list(x = kde$x, y = kde$y, z = kde$z)

    out$dens[ok] <- fields::interp.surface(surf, cbind(xx[ok], yy[ok]))
    out
  }

  spag_d <- spaghetti_df %>%
    add_kde_density(x = "age", y = "phenoage", n = 250) %>%
    group_by(sampleID) %>%
    arrange(age, .by_group = TRUE) %>%
    mutate(
      xend = lead(age),
      yend = lead(phenoage),
      dens_seg = (dens + lead(dens)) / 2
    ) %>%
    ungroup() %>%
    filter(!is.na(xend), !is.na(yend), !is.na(dens_seg))

  phenoage_plot <- phenoage_plot +
    # tiny observation dots first
    geom_point(
      data = spaghetti_df,
      aes(x = age, y = phenoage),
      alpha = 0.50,
      size = 0.10,
      show.legend = FALSE
    ) +
    # density-colored spaghetti segments
    geom_segment(
      data = spag_d,
      aes(x = age, y = phenoage, xend = xend, yend = yend, color = dens_seg),
      linewidth = 0.3,
      alpha = 0.30,
      lineend = "round",
      show.legend = FALSE
    ) +
    guides(color = "none")
}

# Add rotated CI rectangles, lines, and slope labels
phenoage_plot <- phenoage_plot +
  geom_polygon(
    data = polygon_df,
    aes(x = x, y = y, group = group),
    fill = "darkred", alpha = 0.3
  ) +
  geom_segment(
    data = plot_data,
    aes(x = x0, xend = x1, y = y0, yend = y1),
    color = "darkred", size = 1
  ) +
  ggtext::geom_richtext(
    data = plot_data,
    aes(x = label_x, y = label_y, label = label),
    size = 3,
    fontface = "bold",
    color = "black",
    fill = alpha("white", 0.4),
    label.color = NA,
    vjust = -2,
    hjust = 0.6
  ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  coord_fixed(ratio = 1, xlim = c(15, 85), ylim = c(15, 100)) +
  scale_x_continuous(breaks = seq(10, 85, by = 10)) +
  scale_y_continuous(breaks = seq(10, 100, by = 10)) +
  labs(x = "Chronological Age", y = "PhenoAge") +
  theme_minimal()

print(phenoage_plot)
```
```{r}
png("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/PhD_studies/Study_1/Script/Current_version/Longitudinal/trajectory_women_phe_v4_spag_Lukas.png", 
    width = 6, height = 6, units = 'in', res = 1000)
print(phenoage_plot)  # Send plot to file
dev.off()  # Close device
```

## 2.6 Creating figure colored by density (KDM)
```{r}
library(dplyr)
library(ggplot2)
library(stringr)
library(MASS)
library(fields)

# Assume slope_table, to1po2_clean, and fit already exist

# Extract global intercept from model
intercept <- coef(fit)["(Intercept)"]

# Construct CI-based rotated rectangles
plot_data <- slope_table %>%
  rowwise() %>%
  mutate(
    x0 = start,
    x1 = stop,
    y0 = intercept + slope * x0,
    y1 = intercept + slope * x1,

    dx = x1 - x0,
    dy = y1 - y0,
    L = sqrt(dx^2 + dy^2),

    nx = -dy / L,
    ny = dx / L,

    slope_ci = (upper_CI - lower_CI) / 2,
    w = slope_ci * dx / sqrt(1 + slope^2),

    xA = x0 + w * nx,
    yA = y0 + w * ny,
    xB = x1 + w * nx,
    yB = y1 + w * ny,
    xC = x1 - w * nx,
    yC = y1 - w * ny,
    xD = x0 - w * nx,
    yD = y0 - w * ny,

    label_x = (x0 + x1) / 2,
    label_y = (y0 + y1) / 2,
    label = sprintf("%.2f", slope),
    group = paste0("bin_", start)
  ) %>%
  ungroup()

plot_data <- plot_data %>%
  filter(stop <= 82)

# Long-format polygon
polygon_df <- plot_data %>%
  rowwise() %>%
  do(data.frame(
    group = .$group,
    x = c(.$xA, .$xB, .$xC, .$xD),
    y = c(.$yA, .$yB, .$yC, .$yD)
  )) %>%
  ungroup()

# Optional spaghetti overlay toggle
show_spaghetti <- TRUE

# Start ggplot
kdm_plot <- ggplot()

# Add spaghetti first (behind everything)
if (show_spaghetti) {
  set.seed(42)
  subset_ids <- sample(
    unique(to1po2_clean$sampleID),
    floor(0.01 * length(unique(to1po2_clean$sampleID)))
  )

  spaghetti_df <- to1po2_clean %>%
    filter(sampleID %in% subset_ids)

  # Same density approach as your previous KDM plot
  add_kde_density <- function(df, x = "age", y = "kdm", n = 250) {
    xx <- df[[x]]
    yy <- df[[y]]

    ok <- is.finite(xx) & is.finite(yy)

    out <- df
    out$dens <- NA_real_

    if (sum(ok) < 2 || length(unique(xx[ok])) < 2 || length(unique(yy[ok])) < 2) {
      return(out)
    }

    kde <- MASS::kde2d(xx[ok], yy[ok], n = n)
    surf <- list(x = kde$x, y = kde$y, z = kde$z)

    out$dens[ok] <- fields::interp.surface(surf, cbind(xx[ok], yy[ok]))
    out
  }

  spag_d <- spaghetti_df %>%
    add_kde_density(x = "age", y = "kdm", n = 250) %>%
    group_by(sampleID) %>%
    arrange(age, .by_group = TRUE) %>%
    mutate(
      xend = lead(age),
      yend = lead(kdm),
      dens_seg = (dens + lead(dens)) / 2
    ) %>%
    ungroup() %>%
    filter(!is.na(xend), !is.na(yend), !is.na(dens_seg))

  kdm_plot <- kdm_plot +
    # tiny observation dots first
    geom_point(
      data = spaghetti_df,
      aes(x = age, y = kdm),
      alpha = 0.50,
      size = 0.10,
      show.legend = FALSE
    ) +
    # density-colored spaghetti segments
    geom_segment(
      data = spag_d,
      aes(x = age, y = kdm, xend = xend, yend = yend, color = dens_seg),
      linewidth = 0.3,
      alpha = 0.30,
      lineend = "round",
      show.legend = FALSE
    ) +
    guides(color = "none")
}

# Add rotated CI rectangles, lines, and slope labels
kdm_plot <- kdm_plot +
  geom_polygon(
    data = polygon_df,
    aes(x = x, y = y, group = group),
    fill = "darkred", alpha = 0.3
  ) +
  geom_segment(
    data = plot_data,
    aes(x = x0, xend = x1, y = y0, yend = y1),
    color = "darkred", size = 1
  ) +
  ggtext::geom_richtext(
    data = plot_data,
    aes(x = label_x, y = label_y, label = label),
    size = 3,
    fontface = "bold",
    color = "black",
    fill = alpha("white", 0.4),
    label.color = NA,
    vjust = -2,
    hjust = 0.6
  ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  coord_fixed(ratio = 1, xlim = c(15, 85), ylim = c(15, 100)) +
  scale_x_continuous(breaks = seq(10, 85, by = 10)) +
  scale_y_continuous(breaks = seq(10, 100, by = 10)) +
  labs(x = "Chronological Age", y = "KDM") +
  theme_minimal()

print(kdm_plot)

```
```{r}
png("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/PhD_studies/Study_1/Script/Current_version/Longitudinal/trajectory_all_kdm_v4_spag_Lukas.png", 
    width = 6, height = 6, units = 'in', res = 1000)
print(kdm_plot)  # Send plot to file
dev.off()  # Close device
```

# 3.0 Statistical sex differences in trajectories

## 3.1 PhenoAge
```{r}
# Load packages
library(dplyr)
library(sandwich)
library(stringr)

# -----------------------------
# 1. Prepare data
# -----------------------------
to1po2_clean <- results_dt %>%
  filter(!is.na(age), !is.na(phenoage), !is.na(gender)) %>%
  group_by(sampleID) %>%
  filter(n() >= 3) %>%
  ungroup() %>%
  mutate(
    age_bin = cut(
      age,
      breaks = seq(17, ceiling(max(age, na.rm = TRUE)), by = 5),
      include.lowest = TRUE,
      right = FALSE
    ),
    gender = factor(gender, levels = c(1, 2), labels = c("Men", "Women"))
  )

# -----------------------------
# 2. Fit joint model
# -----------------------------
fit <- lm(phenoage ~ age * age_bin * gender, data = to1po2_clean)

# Cluster-robust covariance matrix
vc <- vcovCL(fit, cluster = to1po2_clean$sampleID)

# Coefficients
coefs <- coef(fit)

# Optional: inspect coefficient names if debugging
# print(names(coefs))

# -----------------------------
# 3. Compute slopes and sex differences by age bin
# -----------------------------
z <- 1.96
bins <- levels(to1po2_clean$age_bin)

out <- data.frame(
  age_bin = bins,
  slope_men = NA_real_,
  slope_women = NA_real_,
  diff_women_minus_men = NA_real_,
  lower_CI = NA_real_,
  upper_CI = NA_real_,
  p_value = NA_real_
)

for (i in seq_along(bins)) {
  
  bin <- bins[i]
  
  # Term names
  term_age_bin <- paste0("age:age_bin", bin)
  term_gender  <- "age:genderWomen"
  term_three   <- paste0("age:age_bin", bin, ":genderWomen")
  
  # Base age slope
  beta_age <- coefs["age"]
  
  # Age-bin slope deviation (0 for reference bin)
  beta_bin <- if (term_age_bin %in% names(coefs)) coefs[term_age_bin] else 0
  
  # Women vs men slope difference in reference bin
  beta_gender <- if (term_gender %in% names(coefs)) coefs[term_gender] else 0
  
  # Additional women vs men slope difference in this bin
  beta_three <- if (term_three %in% names(coefs)) coefs[term_three] else 0
  
  # Slopes
  slope_men <- beta_age + beta_bin
  slope_women <- beta_age + beta_bin + beta_gender + beta_three
  
  # Difference: women - men
  diff <- beta_gender + beta_three
  
  # Variance of difference
  if (term_three %in% rownames(vc)) {
    var_diff <- vc[term_gender, term_gender] +
      vc[term_three, term_three] +
      2 * vc[term_gender, term_three]
  } else {
    var_diff <- vc[term_gender, term_gender]
  }
  
  se_diff <- sqrt(var_diff)
  lower <- diff - z * se_diff
  upper <- diff + z * se_diff
  p <- 2 * pnorm(-abs(diff / se_diff))
  
  out[i, ] <- list(
    bin,
    slope_men,
    slope_women,
    diff,
    lower,
    upper,
    p
  )
}

# -----------------------------
# 4. Clean supplement table
# -----------------------------
supp_table <- out %>%
  mutate(
    age_bin = gsub("\\[|\\)", "", age_bin),
    age_bin = gsub(",", "-", age_bin),
    slope_men = round(slope_men, 4),
    slope_women = round(slope_women, 4),
    diff_women_minus_men = round(diff_women_minus_men, 4),
    lower_CI = round(lower_CI, 4),
    upper_CI = round(upper_CI, 4),
    p_value = ifelse(p_value < 0.001, "<0.001", sprintf("%.3f", p_value))
  )

sex_difference_trajectories_phe <- supp_table %>%
  dplyr::rename(
    `Age interval` = age_bin,
    `Slope men` = slope_men,
    `Slope women` = slope_women,
    `Difference (women - men)` = diff_women_minus_men,
    `Lower CI` = lower_CI,
    `Upper CI` = upper_CI,
    `P-value` = p_value
  )

print(sex_difference_trajectories_phe)
```
```{r}
write.table(as.data.frame(sex_difference_trajectories_phe),
            file = "sex_difference_trajectories_phe.csv",
            sep = ";",
            dec = ",",
            row.names = FALSE,
            quote = FALSE,
            fileEncoding = "Latin1") 
```

## 3.2 KDM
```{r}
# Load packages
library(dplyr)
library(sandwich)
library(stringr)

# -----------------------------
# 1. Prepare data
# -----------------------------
to1po2_clean <- results_dt %>%
  filter(!is.na(age), !is.na(kdm), !is.na(gender)) %>%
  group_by(sampleID) %>%
  filter(n() >= 3) %>%
  ungroup() %>%
  mutate(
    age_bin = cut(
      age,
      breaks = seq(17, ceiling(max(age, na.rm = TRUE)), by = 5),
      include.lowest = TRUE,
      right = FALSE
    ),
    gender = factor(gender, levels = c(1, 2), labels = c("Men", "Women"))
  )

# -----------------------------
# 2. Fit joint model
# -----------------------------
fit <- lm(kdm ~ age * age_bin * gender, data = to1po2_clean)

# Cluster-robust covariance matrix
vc <- vcovCL(fit, cluster = to1po2_clean$sampleID)

# Coefficients
coefs <- coef(fit)

# Optional: inspect coefficient names if debugging
# print(names(coefs))

# -----------------------------
# 3. Compute slopes and sex differences by age bin
# -----------------------------
z <- 1.96
bins <- levels(to1po2_clean$age_bin)

out <- data.frame(
  age_bin = bins,
  slope_men = NA_real_,
  slope_women = NA_real_,
  diff_women_minus_men = NA_real_,
  lower_CI = NA_real_,
  upper_CI = NA_real_,
  p_value = NA_real_
)

for (i in seq_along(bins)) {
  
  bin <- bins[i]
  
  # Term names
  term_age_bin <- paste0("age:age_bin", bin)
  term_gender  <- "age:genderWomen"
  term_three   <- paste0("age:age_bin", bin, ":genderWomen")
  
  # Base age slope
  beta_age <- coefs["age"]
  
  # Age-bin slope deviation (0 for reference bin)
  beta_bin <- if (term_age_bin %in% names(coefs)) coefs[term_age_bin] else 0
  
  # Women vs men slope difference in reference bin
  beta_gender <- if (term_gender %in% names(coefs)) coefs[term_gender] else 0
  
  # Additional women vs men slope difference in this bin
  beta_three <- if (term_three %in% names(coefs)) coefs[term_three] else 0
  
  # Slopes
  slope_men <- beta_age + beta_bin
  slope_women <- beta_age + beta_bin + beta_gender + beta_three
  
  # Difference: women - men
  diff <- beta_gender + beta_three
  
  # Variance of difference
  if (term_three %in% rownames(vc)) {
    var_diff <- vc[term_gender, term_gender] +
      vc[term_three, term_three] +
      2 * vc[term_gender, term_three]
  } else {
    var_diff <- vc[term_gender, term_gender]
  }
  
  se_diff <- sqrt(var_diff)
  lower <- diff - z * se_diff
  upper <- diff + z * se_diff
  p <- 2 * pnorm(-abs(diff / se_diff))
  
  out[i, ] <- list(
    bin,
    slope_men,
    slope_women,
    diff,
    lower,
    upper,
    p
  )
}

# -----------------------------
# 4. Clean supplement table
# -----------------------------
supp_table <- out %>%
  mutate(
    age_bin = gsub("\\[|\\)", "", age_bin),
    age_bin = gsub(",", "-", age_bin),
    slope_men = round(slope_men, 4),
    slope_women = round(slope_women, 4),
    diff_women_minus_men = round(diff_women_minus_men, 4),
    lower_CI = round(lower_CI, 4),
    upper_CI = round(upper_CI, 4),
    p_value = ifelse(p_value < 0.001, "<0.001", sprintf("%.3f", p_value))
  )

sex_difference_trajectories_kdm <- supp_table %>%
  dplyr::rename(
    `Age interval` = age_bin,
    `Slope men` = slope_men,
    `Slope women` = slope_women,
    `Difference (women - men)` = diff_women_minus_men,
    `Lower CI` = lower_CI,
    `Upper CI` = upper_CI,
    `P-value` = p_value
  )

print(sex_difference_trajectories_kdm)
```

```{r}
write.table(as.data.frame(sex_difference_trajectories_kdm),
            file = "sex_difference_trajectories_kdm.csv",
            sep = ";",
            dec = ",",
            row.names = FALSE,
            quote = FALSE,
            fileEncoding = "Latin1") 
```
