---
title: "Calculating_BA_B"
output: html_notebook
author: Karolina Gustavsson
date: 2026-04-10
---

Input data: AMORIS

Important note: For KDM we need to train on a uniform distribution for chronological age and sex. In addition, we look at the size of NHANES3 which is the default training set in BioAge to determine the size of the training set for our subset of AMORIS (that we can't use as a training set since we don't have enough overlapping biomarkers). 

# 1.0 Preparing variables

## 1.1 Loading packages
```{r}
options(repos = "http://nexus.ki.se/repository/cran.r-project.org")
.libPaths("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/shun_trial/lib")
source("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/shun_trial/function.R")

library(haven)
library(broom)
library(dplyr)
library(flexsurv)
library(ggplot2)
library(htmlTable)
library(survival)
library(survey)
library(tidyr)
library(rmarkdown)
library(ggplot2)
```

## 1.2 Optional: Determine size of NHANES3 training set in the BioAge package
```{r}
load("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/BioAge-master/BioAge/data/NHANES3.rda")

NH3bm <- c("albumin","alp","lncrp","totchol","lncreat","hba1c","sbp","bun","uap","lymph","mcv","wbc")

NH3 <- NHANES3 %>%
  group_by(gender) %>%
  filter(
    if_else(
      gender == 1,
      across(all_of(NH3bm), ~ .x > mean(.x, na.rm = TRUE) - 5 * sd(.x, na.rm = TRUE) & .x < mean(.x, na.rm = TRUE) + 5 * sd(.x, na.rm = TRUE)),
      across(all_of(NH3bm), ~ .x > mean(.x, na.rm = TRUE) - 5 * sd(.x, na.rm = TRUE) & .x < mean(.x, na.rm = TRUE) + 5 * sd(.x, na.rm = TRUE))
    )
  ) %>%
  ungroup()

complete_cases_count_before <- sum(complete.cases(NHANES3[NH3bm]))
complete_cases_count_after <- sum(complete.cases(NH3[NH3bm]))

# Print the result
print(complete_cases_count_before) #14 780
print(complete_cases_count_after) #14 589

count_df <- NH3[NH3$age >= 30 & NH3$age <= 75 & NH3$pregnant == 0, ]
# Count the number of rows in the filtered dataframe
num_rows <- nrow(count_df)
num_rows #9869 were in the original training set
```

## 1.3 Load AMORIS
```{r}
amoris <- read_sas("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/Data_ADRD/StudyPop.sas7bdat")
```

## 1.4 Optional: for sensitivity analysis of fasting
```{r}
amoris <- amoris %>%
  filter(fastestatus == "F")
```

## 1.5 Age resolution
We need the rounded to whole years to create our test and training set
```{r}
# Calculate the age in years
amoris$age_w <- as.numeric(difftime(amoris$Provtagningsdatum, amoris$FODDATn, units = "days")) / 365.25

# Optionally round down the age if you want the integer part
amoris$age <- floor(amoris$age_w)
```

## 1.6 Removing outliers
```{r}
#select our 17 biomarkers
biomarkers <- c("S_Alb", "S_Krea", "TC", "TG", "S_K", "S_P", "Fe_maet", "S_Urea", "S_LD", "S_Ca", "fS_TIBC", "fS_Jaern", "S_Alp", "S_Urat", "S_FAMN", "S_Hapt", "fS_Gluk")

AMORIS <- amoris %>%
  group_by(Kon) %>%
  filter(
    if_else(
      Kon == 1, 
      across(all_of(biomarkers), ~ .x > mean(.x, na.rm = TRUE) - 5 * sd(.x, na.rm = TRUE) & .x < mean(.x, na.rm = TRUE) + 5 * sd(.x, na.rm = TRUE)),
      across(all_of(biomarkers), ~ .x > mean(.x, na.rm = TRUE) - 5 * sd(.x, na.rm = TRUE) & .x < mean(.x, na.rm = TRUE) + 5 * sd(.x, na.rm = TRUE))
    )
  ) %>%
  ungroup()

#AMORIS <- AMORIS[complete.cases(AMORIS[, biomarkers]), ]

#rm(biomarkers)
```

## 1.7 Variables required by BioAge
```{r}
colnames(AMORIS)[colnames(AMORIS) == "Kon"] <- "gender"
colnames(AMORIS)[colnames(AMORIS) == "Id"] <- "sampleID"
```

## 1.8 Removing those with outcome before baseline
```{r}
AMORIS <- AMORIS %>%
  mutate(status = ifelse(is.na(DODSDATn), 0, 1)) %>%
  mutate(time = as.numeric(difftime(lastDate, firstDate, units = "weeks")) / 4.33)

negative_counts <- sum(AMORIS$time < 0)
print(paste("Number of negative survival times:", negative_counts))

rm(negative_counts)

# Filtering out negative times
AMORIS <- AMORIS %>% filter(time >= 0)
```

## 1.9 Optional: inspecting number of men and women for each age
```{r}
# Assuming AMORIS has been defined and the gender column adjusted as in the previous example
AMORIS$gender_factor <- factor(AMORIS$gender, levels = c("1", "2"), labels = c("Men", "Women"))

# Create the contingency table and convert it to a data frame
age_gender_table <- with(AMORIS, table(Age = age, Gender = AMORIS$gender_factor))
#print(age_gender_table)

df$Age <- as.numeric(rownames(df))
#rownames(df) <- NULL

df <- as.data.frame.matrix(age_gender_table)

# Calculate the minimum of men and women counts for each age
df$Min = apply(df, 1, min)

# Specify the percentage you want to use, e.g., 30%, for the training set
percentage <- 0.3

# Calculate the percentage of Min and use ceiling to round up to the nearest integer
df$PercentageOfMin = ceiling(df$Min * percentage)

# View the final data frame
print(df)
```

## 1.10 Choosing a maximun age interval for the training set, based on KDM assumptions
```{r}
# This becomes a sliding window problem: ie what is the longest age range we can have and still get a training set which is large enough, here is a simple approach

# We can change train_size so that we get enough people in each age
# We want it to be at least as large as NHANES3 when training KDM/HD/Phenoage in the original article
library(dplyr)

# Assuming AMORIS is your initial data frame and chosen_percentage is predefined
chosen_percentage <- 0.3

amoris_summary <- AMORIS %>%
  group_by(age) %>%
  summarise(
    CountMen = sum(gender == "1"),
    CountWomen = sum(gender == "2"),
    MinCount = pmin(CountMen, CountWomen),
    PercentageOfMin = MinCount * chosen_percentage
  ) %>%
  arrange(age)

# Step 2: Create the new data frame
result_df <- amoris_summary %>%
  mutate(A1 = age,
         A2 = NA) %>%
  arrange(age)

# Corrected for loop to find A2
for (i in 1:(nrow(result_df) - 1)) {
  current_min = result_df$PercentageOfMin[i]
  for (j in (i + 1):nrow(result_df)) {
    if (result_df$PercentageOfMin[j] < current_min) {
      result_df$A2[i] = result_df$age[j]
      break
    } else if (j == nrow(result_df)) {
      result_df$A2[i] = result_df$age[nrow(result_df)]
    }
  }
}

result_df <- result_df %>%
  mutate(delta = ifelse(is.na(A2), "no limit", as.character(A2 - A1))) %>%
  arrange(desc(as.numeric(delta)))

# Step 3: Calculate estimated_t_size with the new considerations
result_df <- result_df %>%
  mutate(
    estimated_t_size = case_when(
      delta == "no limit" & age == max(age) ~ "last",
      PercentageOfMin == 0 ~ "no min",
      delta == "no limit" ~ "last age",
      TRUE ~ as.character(as.numeric(delta) * PercentageOfMin)
    )
  ) %>%
  arrange(desc(as.numeric(estimated_t_size)))

# Step 4: Define train_size and tz, filter the rows, and print the filtered data frame
train_size <- 9869 # Assuming train_size is predefined
tz <- train_size / 2

# Convert estimated_t_size to numeric where applicable
result_df <- result_df %>%
  mutate(numeric_estimated_t_size = as.numeric(estimated_t_size))

# Filter rows where estimated_t_size is larger than tz
filtered_df <- result_df %>%
  filter(numeric_estimated_t_size > tz)

# Sort the filtered rows by delta (as numeric)
sorted_filtered_df <- filtered_df %>%
  mutate(numeric_delta = as.numeric(delta)) %>%
  arrange(desc(numeric_delta))

# Print ordered by training set size
print(filtered_df)

# Print based on largest age range
print(sorted_filtered_df)

#take train_size/delta for the longest timespan to, and then that divided by two to get the number of people per year and sex (for instance 80 women per year)

```

## 1.11 Defining our selected biomarkers
```{r}
#over 0.1 separatley for men and women, S_Ca is rounded up, selected in Selecting_biomarkers.rmd
biomarkers = c("S_Alb", "S_Krea", "TC", "TG", "S_K", "S_Urea", "S_LD", "S_Ca", "S_Urat", "S_FAMN", "S_Hapt", "fS_Gluk") 
```

## 1.12 Only using complete cases
```{r}
AMORIS <- AMORIS[complete.cases(AMORIS[, biomarkers]), ]
```

## 1.13 Making our test and training set
```{r}
# Load necessary libraries
library(dplyr)

# Set seed for reproducibility
set.seed(123)

# Define the specific number of men and women for each age you want in the training set
training_sizes <- expand.grid(
  age = 17:79,         # Age from 25 to 75
  gender = c("1", "2"),         # Gender "1" for men and "2" for women, assuming these are the labels used
  stringsAsFactors = FALSE
)
training_sizes$size <- 87  # Specify 100 individuals for each combination of age and gender

# Ensure the sizes do not exceed available data
training_sizes <- training_sizes %>%
  left_join(AMORIS %>% count(age, gender), by = c("age", "gender")) %>%
  mutate(size = pmin(size, n, na.rm = TRUE)) %>%
  select(-n)

# Sample the specified number of individuals for training using a loop instead of slice_sample
selected_training <- list()  # Initialize an empty list to store results
for (i in seq_along(training_sizes$age)) {
  group_subset <- AMORIS %>%
    filter(age == training_sizes$age[i], gender == training_sizes$gender[i])
  
  if (nrow(group_subset) > 0) {
    group_sample <- group_subset %>%
      sample_n(size = training_sizes$size[i])
    selected_training[[i]] <- group_sample
  }
}

# Combine all sampled groups into one data frame for the training set
new_training_set <- bind_rows(selected_training)

# Create the test set with all remaining individuals
new_test_set <- anti_join(AMORIS, new_training_set, by = "sampleID")

# Ensure no overlap in sampleID
if (any(new_training_set$sampleID %in% new_test_set$sampleID)) {
  stop("Error: Overlap detected between training and test sets.")
}

# Check the counts of men and women in each age group for both sets
counts_training <- new_training_set %>%
  group_by(age, gender) %>%
  summarise(count = n(), .groups = 'drop')

counts_test <- new_test_set %>%
  group_by(age, gender) %>%
  summarise(count = n(), .groups = 'drop')

# Print the counts
print("Training Set Counts:")
print(counts_training)
print("Test Set Counts:")
print(counts_test)
```

## 1.14 Visual inspection
```{r}
# Load the ggplot2 package
library(ggplot2)

# Assuming 'amoris' DataFrame is already loaded and 'Kon' is your gender column with 1 for male and 2 for female
# Convert the numeric codes to a factor to improve plot readability
new_training_set$Sex <- factor(new_training_set$gender, levels = c(1, 2), labels = c("Male", "Female"))

# Create overlapping histograms with transparency
b_train_set <- ggplot(new_training_set, aes(x = age, fill = Sex)) +
  geom_histogram(position = "identity", alpha = 0.5, binwidth = 1) +  # binwidth can be adjusted based on data
  scale_fill_manual(values = c("#6BAED6", "#E78AC3")) +  # Colors can be adjusted
  labs(title = "Age Distribution by Sex", x = "Age", y = "Count") +
  theme_minimal()

b_train_set
```

```{r}
png("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/PhD_studies/Study_1/Script/Current_version/Baseline/2025-06-06/b_train_set.png", 
    width = 6, height = 4, units = 'in', res = 600)
print(b_train_set)  # Send plot to file
dev.off()  # Close device
```

```{r}
# Load the ggplot2 package
library(ggplot2)

# Assuming 'amoris' DataFrame is already loaded and 'Kon' is your gender column with 1 for male and 2 for female
# Convert the numeric codes to a factor to improve plot readability
new_test_set$Sex <- factor(new_test_set$gender, levels = c(1, 2), labels = c("Men", "Women"))

# Create overlapping histograms with transparency
a_test_set <- ggplot(new_test_set, aes(x = age, fill = Sex)) +
  geom_histogram(position = "identity", alpha = 0.5, binwidth = 1) +  # binwidth can be adjusted based on data
  scale_fill_manual(values = c("#6BAED6", "#E78AC3")) +  # Colors can be adjusted
  #labs(title = "Age Distribution by Sex", x = "Age", y = "Count") +
  labs(
    x = "Age",
    y = "Count"
  ) +
  theme_minimal()

a_test_set
```

```{r}
png("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/PhD_studies/Study_1/Script/Current_version/Baseline/2025-06-06/a_test_set_pinkblue.png", 
    width = 6, height = 4, units = 'in', res = 600)
print(a_test_set)  # Send plot to file
dev.off()  # Close device
```

## 1.15 Increasing chronological age resolution
```{r}
new_training_set$age <- new_training_set$age_w
new_test_set$age <- new_test_set$age_w
new_training_set$age <- signif(new_training_set$age, digits = 4)
new_test_set$age <- signif(new_test_set$age, digits = 4)
```

# 2.0 BioAge (KDM, PhenoAge)

## 2.1 KDM
```{r}
## KDM
train_f = kdm_calc(new_training_set %>%
                     filter(gender ==2),
                    #filter(age >=  25 & age <= 75, gender ==2),
                   biomarkers, fit = NULL, s_ba2 = NULL)
train_m = kdm_calc(new_training_set %>%
                     filter(gender ==1),
                    #filter(age >= 25 & age <= 75, gender ==1),
                   biomarkers, fit = NULL, s_ba2 = NULL)
kdm_f = kdm_calc(new_test_set %>%
                   filter (gender ==2), #age >= 25 & age <= 75
                 biomarkers,
                 fit = train_f$fit,
                 s_ba2 = train_f$fit$s_ba2)
kdm_m = kdm_calc(new_test_set %>%
                   filter (gender ==1), #age >= 25 & age <= 75
                 biomarkers,
                 fit = train_m$fit,
                 s_ba2 = train_m$fit$s_ba2)
to1po2_kdm = rbind(kdm_f$data,kdm_m$data)

rm(train_f)
rm(train_m)
rm(kdm_f)
rm(kdm_m)
```

## 2.2 Phenoage
```{r}
train = phenoage_calc(new_training_set, biomarkers)
#Project into the AMORIS_2
phenoage = phenoage_calc(new_test_set, biomarkers, fit = train$fit)
#Extract phenoage dataset, trained on 1 projected on 2 = to1po2, previously data
to1po2_phe = (phenoage$data)

rm(phenoage)
rm(train)
```

## 2.3 Merging
```{r}
to1po2 = left_join(new_test_set, to1po2_kdm[, c("sampleID", "kdm", "kdm_advance")], by = "sampleID") %>%
  left_join(., to1po2_phe[, c("sampleID", "phenoage", "phenoage_advance")], by = "sampleID") 
```

## 2.4 Visual inspection: KDM
```{r}
 agevar_ap = c("kdm","phenoage")
 #prepare labels
 label_ap = c("KDM AMORIS",
              "PhenoAge AMORIS")
 #plot age vs bioage
 c_modified_kdm_phe <- plot_ba(to1po2, agevar_ap, label_ap)
 
 print(c_modified_kdm_phe)
```

```{r}
png("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/PhD_studies/Study_1/Script/Current_version/Baseline/2025-06-06/c_modified_kdm_phe_KDMnotKDMage.png", 
    width = 8, height = 4, units = 'in', res = 600)
print(c_modified_kdm_phe)  # Send plot to file
dev.off()  # Close device
```

## 2.5 Scaling
```{r}
# Standardize BioAge measures and add them to the dataframe
to1po2$kdm_sd <- scale(to1po2$kdm)
to1po2$kdm_advance_sd <- scale(to1po2$kdm_advance)
to1po2$phenoage_sd <- scale(to1po2$phenoage)
to1po2$phenoage_advance_sd <- scale(to1po2$phenoage_advance)
```

```{r}
bioage_vars <- c("kdm", "kdm_advance", "phenoage", "phenoage_advance")

# Compute SDs (in years) for each
bioage_sds <- sapply(to1po2[bioage_vars], sd, na.rm = TRUE)

# Create a readable data frame
bioage_scale_df <- data.frame(
  Variable = names(bioage_sds),
  SD_in_years = round(bioage_sds, 2)
)

# Print the result
print(bioage_scale_df)
```

```{r}
write.table(as.data.frame(bioage_scale_df), "bioage_scale_df_table.csv", sep = ";",
            dec = ",",
            row.names = FALSE,
            quote = FALSE,
            fileEncoding = "Latin1")
```

## 2.6 Correlation with BA
```{r}
get_axis_type <- function(labels){
  return(rep("float", length(labels)) %>% setNames(names(labels))) # Create function to generate axis_type variables for BAA plots
}

png("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/PhD_studies/Study_1/Script/Current_version/Baseline/2025-06-06/d_corrplot_good_F.png", 
    width = 6, height = 4.5, units = 'in', res = 600)

plot_baa(data=to1po2,
              agevar = names(c("kdm"="KDM", "kdm_advance"="KDM\nAge\nGap", "phenoage"="PhenoAge","phenoage_advance"="PhenoAge\nAge\nGap", "age"="Chronological\nage")),
              label = c("kdm" = "KDM\n","kdm_advance"="KDM\nAge\nGap",
                        "phenoage" = "PhenoAge",   "phenoage_advance"="PhenoAge\nAge\nGap",
                        "age"="Chronological\nAge"),
              axis_type = c("kdm" = "float",
                            "kdm_advance" = "float",
                            "phenoage" = "float",
                            "phenoage_advance" = "float",
                            "age" = "float"))

```

## 2.7 Saving the sampleIDs for reference (for the longitudinal version)
```{r}
write.table(as.data.frame(to1po2$sampleID),
            "sampleIDs_baseline.csv",
            sep = ";",
            dec = ",",
            row.names = FALSE,
            quote = FALSE,
            fileEncoding = "Latin1")
```

## 2.8 Saving the training set (for the longitudinal version)
```{r}
write.table(as.data.frame(new_training_set), "training_set_baseline.csv", sep = ";",
            dec = ",",
            row.names = FALSE,
            quote = FALSE,
            fileEncoding = "Latin1")
```

## 2.9 Supplemantary density plot: biological age for men and women
```{r}
library(ggplot2)
library(gridExtra)
library(dplyr)

# Convert gender to factor with labels
to1po2 <- to1po2 %>%
  mutate(Sex = factor(gender, levels = c(1, 2), labels = c("Men", "Women")))

# Define hex colors from your original plot
sex_colors <- c("Men" = "#6BAED6", "Women" = "#E78AC3")

# Base theme
base_theme <- theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(size = 9),
    axis.title = element_text(size = 10),
    strip.text = element_blank()
  )
# Age plot
age_plot <- ggplot(to1po2, aes(x = age, fill = Sex, color = Sex)) +
  geom_density(alpha = 0.4, adjust = 2) +
  scale_fill_manual(values = sex_colors) +
  scale_color_manual(values = sex_colors) +
  labs(x = "Chronological Age", y = "Density") +
  base_theme +
  theme(legend.position = "none")

# KDM plot: include axis labels, exclude legend
kdm_plot <- ggplot(to1po2, aes(x = kdm, fill = Sex, color = Sex)) +
  geom_density(alpha = 0.4, adjust = 2) +
  scale_fill_manual(values = sex_colors) +
  scale_color_manual(values = sex_colors) +
  labs(x = "KDM", y = "Density") +
  base_theme +
  theme(legend.position = "none")

# PhenoAge plot: includes legend
pheno_plot <- ggplot(to1po2, aes(x = phenoage, fill = Sex, color = Sex)) +
  geom_density(alpha = 0.4, adjust = 2) +
  scale_fill_manual(values = sex_colors, name = "Sex") +
  scale_color_manual(values = sex_colors, name = "Sex") +
  labs(x = "PhenoAge", y = "Density") +
  base_theme +
  theme(legend.position = "none")
  
  #theme(legend.title = element_text(face = "bold")) +
  #theme(legend.position = "bottom")  # or "right", "top", "left", or use `c(x, y)` for exact placement

# Create dummy plot just to extract legend
dummy <- ggplot(to1po2, aes(x = phenoage, fill = Sex, color = Sex)) +
  geom_density(alpha = 0.4) +
  scale_fill_manual(values = sex_colors, name = "Sex") +
  scale_color_manual(values = sex_colors, name = "Sex") +
  theme_void() +
  theme(legend.position = "bottom")

# Extract legend (fix: use gtable::gtable_filter)
get_legend <- function(plot) {
  tmp <- ggplotGrob(plot)
  gtable::gtable_filter(tmp, "guide-box")
}

legend_grob <- get_legend(dummy)


# Arrange plots and legend
grid.arrange(
  arrangeGrob(age_plot, kdm_plot, pheno_plot, ncol = 1),
  legend_grob,
  ncol = 1,
  heights = c(3, 0.1)  # adjust as needed
)

# Combine side by side
#grid.arrange(age_plot, kdm_plot, pheno_plot, ncol = 3)
```

```{r}
# Open device BEFORE running grid.arrange
png("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/PhD_studies/Study_1/Script/Current_version/Baseline/2025-06-06/Density_ba_sex_wCA.png",
    width = 3, height = 5, units = "in", res = 600) #with only phe and kdm width = 8, height = 4, units = "in"

# Draw the combined plot (e.g., two-panel)
#gridExtra::grid.arrange(age_plot, kdm_plot, pheno_plot, ncol = 1)
# Arrange plots and legend
grid.arrange(
  arrangeGrob(age_plot, kdm_plot, pheno_plot, ncol = 1),
  legend_grob,
   ncol = 1,
  heights = c(3, 0.3)  # adjust as needed
)

# Close the device
dev.off()
```
