---
title: "CoxPH_B"
output: html_notebook
author: Karolina Gustavsson
date: 2026-04-10
---

Input data: to1po2 from Calculating_BA.rmd
Covariates: sex, education level, income quartiles

# 1.0 Preparing variables

## 1.1 Linking SveDem 
```{r}
SveDem <- read_sas("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/Data_ADRD/SveDem.sas7bdat")
```

```{r}
#ADRD categories
to1po2$S_ADRD <- ifelse(to1po2$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(to1po2$sampleID, SveDem$Id)] %in% c(1,2,3,4,5,6,7,8,9), 1, 0) #1-9
to1po2$S_e_alz <- ifelse(to1po2$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(to1po2$sampleID, SveDem$Id)] == 1, 1, 0) #1
to1po2$S_l_alz <- ifelse(to1po2$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(to1po2$sampleID, SveDem$Id)] == 2, 1, 0) #2
to1po2$S_alz_vas <- ifelse(to1po2$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(to1po2$sampleID, SveDem$Id)] == 3, 1, 0) #3
to1po2$S_vas <- ifelse(to1po2$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(to1po2$sampleID, SveDem$Id)] == 4, 1, 0) #4
to1po2$S_lew <- ifelse(to1po2$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(to1po2$sampleID, SveDem$Id)] == 5, 1, 0) #5
to1po2$S_fro <- ifelse(to1po2$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(to1po2$sampleID, SveDem$Id)] == 6, 1, 0) #6
to1po2$S_par <- ifelse(to1po2$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(to1po2$sampleID, SveDem$Id)] == 7, 1, 0) #7
to1po2$S_uns <- ifelse(to1po2$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(to1po2$sampleID, SveDem$Id)] == 8, 1, 0) #8 
to1po2$S_oth <- ifelse(to1po2$sampleID %in% SveDem$Id & SveDem$DIAGNOSIS_BR[match(to1po2$sampleID, SveDem$Id)] == 9, 1, 0) #9  

# Create the new column by matching sampleID from to1po2 to Id in SveDem
to1po2$S_INDATUM <- SveDem$DIAGNOSIS_DATE_BR[match(to1po2$sampleID, SveDem$Id)]

# ICD codes for the ones not in SveDem
alzheimers_codes <- c(
  "G308", "G309", "G301", "F009", "F002", "G300", "F001", "F000",
  "290,1", "304", "290", "290A", "F00", "305", "290B", "G30", "311A"
)

vascular_codes <- c(
  "293", "290E", "F01", "293,1", "F019", "F012", "F018",
  "F013", "F011", "F010", "F0180", "F0183"
)

other_codes <- c(
  "F039", "F23", "F020", "F024", "F028", "F022", "F03-p", "F021", "F023", "F03-P",
  "290,0", "293,9", "293,2", "293,5", "293,3", "293,4", "F0013", "F0010",
  "306", "290X", "F02", "290W", "F03", "294B", "G311",
  "311B", "G318A", "311C", "F051", "311X"
)

all_codes <- unique(c(alzheimers_codes, vascular_codes, other_codes))

#Then we make an overall category
to1po2$ADRD <- ifelse(
  (to1po2$Dementia != "" & !is.na(to1po2$Dementia)) |
  (to1po2$ATC != "" & !is.na(to1po2$ATC)) |
  (to1po2$S_ADRD == 1),
  1,
  0
)

# Prioritize SveDem info, fallback to Dementia codes
to1po2$Alzheimers <- ifelse(
  to1po2$S_l_alz == 1 | to1po2$S_e_alz == 1 | to1po2$S_alz_vas == 1,
  1,
  ifelse(to1po2$Dementia %in% alzheimers_codes, 1, 0)
)

to1po2$Vascular <- ifelse(
  to1po2$S_vas == 1,
  1,
  ifelse(to1po2$Dementia %in% vascular_codes, 1, 0)
)

to1po2$Other <- ifelse(
  to1po2$S_oth == 1 | to1po2$S_uns == 1 | to1po2$S_par == 1 | to1po2$S_fro == 1 | to1po2$S_lew == 1,
  1,
  ifelse(to1po2$Dementia %in% other_codes, 1, 0)
)

# Adds the people with no dementia diagnosis but ATC code to alzheimers, the people with dementia drugs have alzheimers
to1po2$Alzheimers <- ifelse(to1po2$ATC != "" & to1po2$Dementia == "" & to1po2$S_ADRD == "", 1, to1po2$Alzheimers)
```

## 1.2 Creating income quartiles
```{r}
# Income level as quartiles
# Step 1: Create quartile columns
to1po2$ArbInk_quartile <- cut(
  to1po2$ArbInk,
  breaks = quantile(to1po2$ArbInk, probs = seq(0, 1, 0.25), na.rm = TRUE),
  include.lowest = TRUE,
  labels = 1:4
)

to1po2$DispInk_quartile <- cut(
  to1po2$DispInk,
  breaks = quantile(to1po2$DispInk, probs = seq(0, 1, 0.25), na.rm = TRUE),
  include.lowest = TRUE,
  labels = 1:4
)
# Step 2: Create income_bins with NA if neither value exists
to1po2$income_bins <- ifelse(
  !is.na(to1po2$DispInk_quartile),  # If DispInk_quartile is available, use it
  as.numeric(as.character(to1po2$DispInk_quartile)),  # Convert factor to numeric
  ifelse(
    !is.na(to1po2$ArbInk_quartile),  # Otherwise, if ArbInk_quartile is available, use it
    as.numeric(as.character(to1po2$ArbInk_quartile)),
    NA  # If neither is available, assign NA
  )
)
```

## 1.3 Education level 
```{r}
# Education level
to1po2$NewUtbNiva <- ifelse(
  !is.na(to1po2$Sun2000niva_old) & to1po2$Sun2000niva_old != "*" & !is.na(as.numeric(to1po2$Sun2000niva_old)), # Check if SUN2000_old is numeric and not '*'
  as.numeric(to1po2$Sun2000niva_old),                                                                  # Use SUN2000_old if valid
  ifelse(
    !is.na(to1po2$UtbNiva) & to1po2$UtbNiva != "*" & !is.na(as.numeric(to1po2$UtbNiva)),           # Check if UtbNiva is numeric and not '*'
    as.numeric(to1po2$UtbNiva),                                                                    # Use UtbNiva as fallback
    NA                                                                                             # Otherwise, assign NA
  )
)
```

## 1.4 Age at first date and age at last date
```{r}
# Calculate age at FirstDate in years
to1po2$AgeAtFirstDate <- as.numeric(difftime(to1po2$firstDate, to1po2$FODDATn, units = "days") / 365.25)

### We want to check if lastDate is FDATUM for the people with nonempty ATC col (chr string) who are empty in Dementia (no chr string) and S_ADRD

# Start with the default: age at lastDate
to1po2$AgeAtLastDate <- as.numeric(difftime(to1po2$lastDate, to1po2$FODDATn, units = "days")) / 365.25

# Identify the "meds-only" group (people missed by other sources)
meds_only <- to1po2$ATC != "" & !is.na(to1po2$ATC) &
             (to1po2$Dementia == "" | is.na(to1po2$Dementia)) &
             to1po2$S_ADRD == 0

# For this group, recalculate age using FDATUM
to1po2$AgeAtLastDate[meds_only] <- as.numeric(difftime(to1po2$FDATUM[meds_only], to1po2$FODDATn[meds_only], units = "days")) / 365.25

### We want whichever is first SveDem INDATUM or regular INDATUM

# Step 3: For everyone else – if either INDATUM or S_INDATUM exists, use the earliest one
earliest_dx <- pmin(to1po2$INDATUM, to1po2$S_INDATUM, na.rm = TRUE)

has_dx_date <- !is.na(earliest_dx)

to1po2$AgeAtLastDate[has_dx_date] <- as.numeric(
  difftime(earliest_dx[has_dx_date], to1po2$FODDATn[has_dx_date], units = "days")
) / 365.25
```

## 1.5 Birth cohort
```{r}
#Decades (but first and last is 20 years)
# Extract decade from birth year
to1po2$birthcohort <- paste0(substr(format(to1po2$FODDATn, "%Y"), 1, 3), "0s")

# Manually adjust the first and last groups
to1po2$birthcohort[to1po2$birthcohort %in% c("1890s", "1900s","1910s")] <- "1890s-1910s"
to1po2$birthcohort[to1po2$birthcohort %in% c("1960s","1970s", "1980s", "1990s")] <- "1960s-1990s"

# Check the results
table(to1po2$birthcohort)
```

## 1.6 Function for calculating beta
```{r}
calc_standardized_beta <- function(model, data, covariate) {
    beta <- coef(model)[covariate]
    sd_covariate <- sd(data[[covariate]], na.rm = TRUE)
    std_beta <- beta * sd_covariate
    return(std_beta)
}
```

# 2.0 Time to evenet analysis: main

## 2.1 ADRD
```{r}
# BA measures to be tested
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Correct the initialization of HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),  # Repeat each BA measure twice
  outcome = rep("ADRD", times = length(BA_measures) * 2),
  model = rep(c("Sex-adjusted model", "Multivariable model"), times = length(BA_measures)),
  beta = NA, se = NA, HR = NA, CI_lower = NA, CI_upper = NA, p = NA, p_PH_BA = NA, global_p_PH_BA = NA,
  concordance = NA, N = NA, events = NA , std_beta = NA  # Add a column for the concordance index
)

for (j in 1:length(BA_measures)) {
  # Sex-adjusted model
  fit_gender <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, ADRD) ~", BA_measures[j], "+ gender")), data = to1po2)
  coef_gender <- summary(fit_gender)
  row_index_gender <- (j * 2) - 1
  std_beta_gender <- calc_standardized_beta(fit_gender, to1po2, BA_measures[j])
  HR_BA[row_index_gender, c(4, 5, 9, 12)] <- c(coef_gender$coefficients[1, c(1, 3, 5)], coef_gender$concordance[1])
  HR_BA[row_index_gender, c("HR", "CI_lower", "CI_upper")] <- coef_gender$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_gender, 10] <- cox.zph(fit_gender)$table[1, 3]
  HR_BA[row_index_gender, 11] <- cox.zph(fit_gender)$table["GLOBAL", "p"]
  HR_BA[row_index_gender, "N"] <- coef_gender$n
  HR_BA[row_index_gender, "std_beta"] <- std_beta_gender
  HR_BA[row_index_gender, "events"] <- fit_gender$nevent
  
  # Multivariable model
  fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, ADRD) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva  ")), data = to1po2)
  coef_multivariable <- summary(fit_multivariable)
  row_index_multivariable <- j * 2
  std_beta_multivariable <- calc_standardized_beta(fit_multivariable, to1po2, BA_measures[j])
  HR_BA[row_index_multivariable, c(4, 5, 9, 12)] <- c(coef_multivariable$coefficients[1, c(1, 3, 5)], coef_multivariable$concordance[1])
  HR_BA[row_index_multivariable, c("HR", "CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_multivariable, 10] <- cox.zph(fit_multivariable)$table[1, 3]
  HR_BA[row_index_multivariable, 11] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
  HR_BA[row_index_multivariable, "N"] <- coef_multivariable$n
  HR_BA[row_index_multivariable, "std_beta"] <- std_beta_multivariable
  HR_BA[row_index_multivariable, "events"] <- fit_multivariable$nevent
}

ADRD_results <- HR_BA

# Print the modified HR_BA_adrd data frame
print(HR_BA)
```

## 2.2 Alzheimers
```{r}
# BA measures to be tested
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Correct the initialization of HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),  # Repeat each BA measure twice
  outcome = rep("Alzheimers", times = length(BA_measures) * 2),
  model = rep(c("Sex-adjusted model", "Multivariable model"), times = length(BA_measures)),
  beta = NA, se = NA, HR = NA, CI_lower = NA, CI_upper = NA, p = NA, p_PH_BA = NA, global_p_PH_BA = NA,
  concordance = NA, N = NA, events = NA , std_beta = NA  # Add a column for the concordance index
)

for (j in 1:length(BA_measures)) {
  # Sex-adjusted model
  fit_gender <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, Alzheimers) ~", BA_measures[j], "+ gender")), data = to1po2)
  coef_gender <- summary(fit_gender)
  row_index_gender <- (j * 2) - 1
  std_beta_gender <- calc_standardized_beta(fit_gender, to1po2, BA_measures[j])
  HR_BA[row_index_gender, c(4, 5, 9, 12)] <- c(coef_gender$coefficients[1, c(1, 3, 5)], coef_gender$concordance[1])
  HR_BA[row_index_gender, c("HR", "CI_lower", "CI_upper")] <- coef_gender$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_gender, 10] <- cox.zph(fit_gender)$table[1, 3]
  HR_BA[row_index_gender, 11] <- cox.zph(fit_gender)$table["GLOBAL", "p"]
  HR_BA[row_index_gender, "N"] <- coef_gender$n
  HR_BA[row_index_gender, "std_beta"] <- std_beta_gender
  HR_BA[row_index_gender, "events"] <- fit_gender$nevent
  
  # Multivariable model
  fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, Alzheimers) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva  ")), data = to1po2)
  coef_multivariable <- summary(fit_multivariable)
  row_index_multivariable <- j * 2
  std_beta_multivariable <- calc_standardized_beta(fit_multivariable, to1po2, BA_measures[j])
  HR_BA[row_index_multivariable, c(4, 5, 9, 12)] <- c(coef_multivariable$coefficients[1, c(1, 3, 5)], coef_multivariable$concordance[1])
  HR_BA[row_index_multivariable, c("HR", "CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_multivariable, 10] <- cox.zph(fit_multivariable)$table[1, 3]
  HR_BA[row_index_multivariable, 11] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
  HR_BA[row_index_multivariable, "N"] <- coef_multivariable$n
  HR_BA[row_index_multivariable, "std_beta"] <- std_beta_multivariable
  HR_BA[row_index_multivariable, "events"] <- fit_multivariable$nevent
}

Alzh_results <- HR_BA

# Print the modified HR_BA_adrd data frame
print(HR_BA)
```

## 2.3 Vascular
```{r}
# BA measures to be tested
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Correct the initialization of HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),  # Repeat each BA measure twice
  outcome = rep("Vascular", times = length(BA_measures) * 2),
  model = rep(c("Sex-adjusted model", "Multivariable model"), times = length(BA_measures)),
  beta = NA, se = NA, HR = NA, CI_lower = NA, CI_upper = NA, p = NA, p_PH_BA = NA, global_p_PH_BA = NA,
  concordance = NA, N = NA, events = NA , std_beta = NA  # Add a column for the concordance index
)

for (j in 1:length(BA_measures)) {
  # Sex-adjusted model
  fit_gender <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, Vascular) ~", BA_measures[j], "+ gender")), data = to1po2)
  coef_gender <- summary(fit_gender)
  row_index_gender <- (j * 2) - 1
  std_beta_gender <- calc_standardized_beta(fit_gender, to1po2, BA_measures[j])
  HR_BA[row_index_gender, c(4, 5, 9, 12)] <- c(coef_gender$coefficients[1, c(1, 3, 5)], coef_gender$concordance[1])
  HR_BA[row_index_gender, c("HR", "CI_lower", "CI_upper")] <- coef_gender$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_gender, 10] <- cox.zph(fit_gender)$table[1, 3]
  HR_BA[row_index_gender, 11] <- cox.zph(fit_gender)$table["GLOBAL", "p"]
  HR_BA[row_index_gender, "N"] <- coef_gender$n
  HR_BA[row_index_gender, "std_beta"] <- std_beta_gender
  HR_BA[row_index_gender, "events"] <- fit_gender$nevent
  
  # Multivariable model
  fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, Vascular) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva   ")), data = to1po2)
  coef_multivariable <- summary(fit_multivariable)
  row_index_multivariable <- j * 2
  std_beta_multivariable <- calc_standardized_beta(fit_multivariable, to1po2, BA_measures[j])
  HR_BA[row_index_multivariable, c(4, 5, 9, 12)] <- c(coef_multivariable$coefficients[1, c(1, 3, 5)], coef_multivariable$concordance[1])
  HR_BA[row_index_multivariable, c("HR", "CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_multivariable, 10] <- cox.zph(fit_multivariable)$table[1, 3]
  HR_BA[row_index_multivariable, 11] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
  HR_BA[row_index_multivariable, "N"] <- coef_multivariable$n
  HR_BA[row_index_multivariable, "std_beta"] <- std_beta_multivariable
  HR_BA[row_index_multivariable, "events"] <- fit_multivariable$nevent
}

Vasc_results <- HR_BA

# Print the modified HR_BA_adrd data frame
print(HR_BA)
```

## 2.4 Other
```{r}
# BA measures to be tested
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Correct the initialization of HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),  # Repeat each BA measure twice
  outcome = rep("Other", times = length(BA_measures) * 2),
  model = rep(c("Sex-adjusted model", "Multivariable model"), times = length(BA_measures)),
  beta = NA, se = NA, HR = NA, CI_lower = NA, CI_upper = NA, p = NA, p_PH_BA = NA, global_p_PH_BA = NA,
  concordance = NA, N = NA, events = NA , std_beta = NA  # Add a column for the concordance index
)

for (j in 1:length(BA_measures)) {
  # Sex-adjusted model
  fit_gender <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, Other) ~", BA_measures[j], "+ gender")), data = to1po2)
  coef_gender <- summary(fit_gender)
  row_index_gender <- (j * 2) - 1
  std_beta_gender <- calc_standardized_beta(fit_gender, to1po2, BA_measures[j])
  HR_BA[row_index_gender, c(4, 5, 9, 12)] <- c(coef_gender$coefficients[1, c(1, 3, 5)], coef_gender$concordance[1])
  HR_BA[row_index_gender, c("HR", "CI_lower", "CI_upper")] <- coef_gender$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_gender, 10] <- cox.zph(fit_gender)$table[1, 3]
  HR_BA[row_index_gender, 11] <- cox.zph(fit_gender)$table["GLOBAL", "p"]
  HR_BA[row_index_gender, "N"] <- coef_gender$n
  HR_BA[row_index_gender, "std_beta"] <- std_beta_gender
  HR_BA[row_index_gender, "events"] <- fit_gender$nevent
  
  # Multivariable model
  fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, Other) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva   ")), data = to1po2)
  coef_multivariable <- summary(fit_multivariable)
  row_index_multivariable <- j * 2
  std_beta_multivariable <- calc_standardized_beta(fit_multivariable, to1po2, BA_measures[j])
  HR_BA[row_index_multivariable, c(4, 5, 9, 12)] <- c(coef_multivariable$coefficients[1, c(1, 3, 5)], coef_multivariable$concordance[1])
  HR_BA[row_index_multivariable, c("HR", "CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_multivariable, 10] <- cox.zph(fit_multivariable)$table[1, 3]
  HR_BA[row_index_multivariable, 11] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
  HR_BA[row_index_multivariable, "N"] <- coef_multivariable$n
  HR_BA[row_index_multivariable, "std_beta"] <- std_beta_multivariable
  HR_BA[row_index_multivariable, "events"] <- fit_multivariable$nevent
}

Othe_results <- HR_BA

# Print the modified HR_BA_adrd data frame
print(HR_BA)
```

## 2.5 Binding and saving 
```{r}
# Combine all data frames into one large data frame
combined_results <- rbind(
  Alzh_results,
  ADRD_results,
  Othe_results,
  Vasc_results
)

```

```{r}
write.table(as.data.frame(combined_results), "combined_results_ADRD_main.csv", sep = ";",
            dec = ",",
            row.names = FALSE,
            quote = FALSE,
            fileEncoding = "Latin1")
```

# 3.0 Time to evenet analysis: SveDem

## 3.1 Only SveDem ADRD cases, removing cases from other registers
```{r}
# 1. Keep people with no dementia info
no_dementia <- to1po2[is.na(to1po2$Dementia) | to1po2$Dementia == "", ]

# 2. Keep everyone who is in SveDem (regardless of Dementia info)
in_svedem <- to1po2[to1po2$sampleID %in% SveDem$Id, ]

# 3. Combine
to1po2_SveDem_D <- rbind(no_dementia, in_svedem)

to1po2_SveDem_D <- unique(to1po2_SveDem_D) #247 254 before with incorrect way, now the same number

```

## 3.2 ADRD (SveDem)
```{r}
# BA measures to be tested
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Correct the initialization of HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),  # Repeat each BA measure twice
  outcome = rep("S_ADRD", times = length(BA_measures) * 2),
  model = rep(c("Sex-adjusted model", "Multivariable model"), times = length(BA_measures)),
  beta = NA, se = NA, HR = NA, CI_lower = NA, CI_upper = NA, p = NA, p_PH_BA = NA, global_p_PH_BA = NA,
  concordance = NA, N = NA, events = NA , std_beta = NA  # Add a column for the concordance index
)

for (j in 1:length(BA_measures)) {
  # Sex-adjusted model
  fit_gender <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_ADRD) ~", BA_measures[j], "+ gender")), data = to1po2_SveDem_D)
  coef_gender <- summary(fit_gender)
  row_index_gender <- (j * 2) - 1
  std_beta_gender <- calc_standardized_beta(fit_gender, to1po2_SveDem_D, BA_measures[j])
  HR_BA[row_index_gender, c(4, 5, 9, 12)] <- c(coef_gender$coefficients[1, c(1, 3, 5)], coef_gender$concordance[1])
  HR_BA[row_index_gender, c("HR", "CI_lower", "CI_upper")] <- coef_gender$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_gender, 10] <- cox.zph(fit_gender)$table[1, 3]
  HR_BA[row_index_gender, 11] <- cox.zph(fit_gender)$table["GLOBAL", "p"]
  HR_BA[row_index_gender, "N"] <- coef_gender$n
  HR_BA[row_index_gender, "std_beta"] <- std_beta_gender
  HR_BA[row_index_gender, "events"] <- fit_gender$nevent
  
  # Multivariable model
  fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_ADRD) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva   ")), data = to1po2_SveDem_D)
  coef_multivariable <- summary(fit_multivariable)
  row_index_multivariable <- j * 2
  std_beta_multivariable <- calc_standardized_beta(fit_multivariable, to1po2_SveDem_D, BA_measures[j])
  HR_BA[row_index_multivariable, c(4, 5, 9, 12)] <- c(coef_multivariable$coefficients[1, c(1, 3, 5)], coef_multivariable$concordance[1])
  HR_BA[row_index_multivariable, c("HR", "CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_multivariable, 10] <- cox.zph(fit_multivariable)$table[1, 3]
  HR_BA[row_index_multivariable, 11] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
  HR_BA[row_index_multivariable, "N"] <- coef_multivariable$n
  HR_BA[row_index_multivariable, "std_beta"] <- std_beta_multivariable
  HR_BA[row_index_multivariable, "events"] <- fit_multivariable$nevent
}

S_ADRD_results <- HR_BA

# Print the modified HR_BA_adrd data frame
print(HR_BA)
```

## 3.3 Early alzheimers (SveDem)
```{r}
# BA measures to be tested
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Correct the initialization of HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),  # Repeat each BA measure twice
  outcome = rep("S_e_alz", times = length(BA_measures) * 2),
  model = rep(c("Sex-adjusted model", "Multivariable model"), times = length(BA_measures)),
  beta = NA, se = NA, HR = NA, CI_lower = NA, CI_upper = NA, p = NA, p_PH_BA = NA, global_p_PH_BA = NA,
  concordance = NA, N = NA, events = NA , std_beta = NA  # Add a column for the concordance index
)

for (j in 1:length(BA_measures)) {
  # Sex-adjusted model
  fit_gender <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_e_alz) ~", BA_measures[j], "+ gender")), data = to1po2_SveDem_D)
  coef_gender <- summary(fit_gender)
  row_index_gender <- (j * 2) - 1
  std_beta_gender <- calc_standardized_beta(fit_gender, to1po2_SveDem_D, BA_measures[j])
  HR_BA[row_index_gender, c(4, 5, 9, 12)] <- c(coef_gender$coefficients[1, c(1, 3, 5)], coef_gender$concordance[1])
  HR_BA[row_index_gender, c("HR", "CI_lower", "CI_upper")] <- coef_gender$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_gender, 10] <- cox.zph(fit_gender)$table[1, 3]
  HR_BA[row_index_gender, 11] <- cox.zph(fit_gender)$table["GLOBAL", "p"]
  HR_BA[row_index_gender, "N"] <- coef_gender$n
  HR_BA[row_index_gender, "std_beta"] <- std_beta_gender
  HR_BA[row_index_gender, "events"] <- fit_gender$nevent
  
  # Multivariable model
  fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_e_alz) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva   ")), data = to1po2_SveDem_D)
  coef_multivariable <- summary(fit_multivariable)
  row_index_multivariable <- j * 2
  std_beta_multivariable <- calc_standardized_beta(fit_multivariable, to1po2_SveDem_D, BA_measures[j])
  HR_BA[row_index_multivariable, c(4, 5, 9, 12)] <- c(coef_multivariable$coefficients[1, c(1, 3, 5)], coef_multivariable$concordance[1])
  HR_BA[row_index_multivariable, c("HR", "CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_multivariable, 10] <- cox.zph(fit_multivariable)$table[1, 3]
  HR_BA[row_index_multivariable, 11] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
  HR_BA[row_index_multivariable, "N"] <- coef_multivariable$n
  HR_BA[row_index_multivariable, "std_beta"] <- std_beta_multivariable
  HR_BA[row_index_multivariable, "events"] <- fit_multivariable$nevent
}

S_e_alz_results <- HR_BA

# Print the modified HR_BA_adrd data frame
print(HR_BA)
```

## 3.4 Late alzheimers (SveDem)
```{r}
# BA measures to be tested
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Correct the initialization of HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),  # Repeat each BA measure twice
  outcome = rep("S_l_alz", times = length(BA_measures) * 2),
  model = rep(c("Sex-adjusted model", "Multivariable model"), times = length(BA_measures)),
  beta = NA, se = NA, HR = NA, CI_lower = NA, CI_upper = NA, p = NA, p_PH_BA = NA, global_p_PH_BA = NA,
  concordance = NA, N = NA, events = NA , std_beta = NA  # Add a column for the concordance index
)

for (j in 1:length(BA_measures)) {
  # Sex-adjusted model
  fit_gender <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_l_alz) ~", BA_measures[j], "+ gender")), data = to1po2_SveDem_D)
  coef_gender <- summary(fit_gender)
  row_index_gender <- (j * 2) - 1
  std_beta_gender <- calc_standardized_beta(fit_gender, to1po2_SveDem_D, BA_measures[j])
  HR_BA[row_index_gender, c(4, 5, 9, 12)] <- c(coef_gender$coefficients[1, c(1, 3, 5)], coef_gender$concordance[1])
  HR_BA[row_index_gender, c("HR", "CI_lower", "CI_upper")] <- coef_gender$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_gender, 10] <- cox.zph(fit_gender)$table[1, 3]
  HR_BA[row_index_gender, 11] <- cox.zph(fit_gender)$table["GLOBAL", "p"]
  HR_BA[row_index_gender, "N"] <- coef_gender$n
  HR_BA[row_index_gender, "std_beta"] <- std_beta_gender
  HR_BA[row_index_gender, "events"] <- fit_gender$nevent
  
  # Multivariable model
  fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_l_alz) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva   ")), data = to1po2_SveDem_D)
  coef_multivariable <- summary(fit_multivariable)
  row_index_multivariable <- j * 2
  std_beta_multivariable <- calc_standardized_beta(fit_multivariable, to1po2_SveDem_D, BA_measures[j])
  HR_BA[row_index_multivariable, c(4, 5, 9, 12)] <- c(coef_multivariable$coefficients[1, c(1, 3, 5)], coef_multivariable$concordance[1])
  HR_BA[row_index_multivariable, c("HR", "CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_multivariable, 10] <- cox.zph(fit_multivariable)$table[1, 3]
  HR_BA[row_index_multivariable, 11] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
  HR_BA[row_index_multivariable, "N"] <- coef_multivariable$n
  HR_BA[row_index_multivariable, "std_beta"] <- std_beta_multivariable
  HR_BA[row_index_multivariable, "events"] <- fit_multivariable$nevent
}

S_l_alz_results <- HR_BA

# Print the modified HR_BA_adrd data frame
print(HR_BA)
```

## 3.5 Mixed (alzheimers and vascular, SveDem)
```{r}
# BA measures to be tested
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Correct the initialization of HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),  # Repeat each BA measure twice
  outcome = rep("S_alz_vas", times = length(BA_measures) * 2),
  model = rep(c("Sex-adjusted model", "Multivariable model"), times = length(BA_measures)),
  beta = NA, se = NA, HR = NA, CI_lower = NA, CI_upper = NA, p = NA, p_PH_BA = NA, global_p_PH_BA = NA,
  concordance = NA, N = NA, events = NA , std_beta = NA  # Add a column for the concordance index
)

for (j in 1:length(BA_measures)) {
  # Sex-adjusted model
  fit_gender <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_alz_vas) ~", BA_measures[j], "+ gender")), data = to1po2_SveDem_D)
  coef_gender <- summary(fit_gender)
  row_index_gender <- (j * 2) - 1
  std_beta_gender <- calc_standardized_beta(fit_gender, to1po2_SveDem_D, BA_measures[j])
  HR_BA[row_index_gender, c(4, 5, 9, 12)] <- c(coef_gender$coefficients[1, c(1, 3, 5)], coef_gender$concordance[1])
  HR_BA[row_index_gender, c("HR", "CI_lower", "CI_upper")] <- coef_gender$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_gender, 10] <- cox.zph(fit_gender)$table[1, 3]
  HR_BA[row_index_gender, 11] <- cox.zph(fit_gender)$table["GLOBAL", "p"]
  HR_BA[row_index_gender, "N"] <- coef_gender$n
  HR_BA[row_index_gender, "std_beta"] <- std_beta_gender
  HR_BA[row_index_gender, "events"] <- fit_gender$nevent
  
  # Multivariable model
  fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_alz_vas) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva   ")), data = to1po2_SveDem_D)
  coef_multivariable <- summary(fit_multivariable)
  row_index_multivariable <- j * 2
  std_beta_multivariable <- calc_standardized_beta(fit_multivariable, to1po2_SveDem_D, BA_measures[j])
  HR_BA[row_index_multivariable, c(4, 5, 9, 12)] <- c(coef_multivariable$coefficients[1, c(1, 3, 5)], coef_multivariable$concordance[1])
  HR_BA[row_index_multivariable, c("HR", "CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_multivariable, 10] <- cox.zph(fit_multivariable)$table[1, 3]
  HR_BA[row_index_multivariable, 11] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
  HR_BA[row_index_multivariable, "N"] <- coef_multivariable$n
  HR_BA[row_index_multivariable, "std_beta"] <- std_beta_multivariable
  HR_BA[row_index_multivariable, "events"] <- fit_multivariable$nevent
}

S_alz_vas_results <- HR_BA

# Print the modified HR_BA_adrd data frame
print(HR_BA)
```

## 3.6 Vascular (SveDem)
```{r}
# BA measures to be tested
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Correct the initialization of HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),  # Repeat each BA measure twice
  outcome = rep("S_vas", times = length(BA_measures) * 2),
  model = rep(c("Sex-adjusted model", "Multivariable model"), times = length(BA_measures)),
  beta = NA, se = NA, HR = NA, CI_lower = NA, CI_upper = NA, p = NA, p_PH_BA = NA, global_p_PH_BA = NA,
  concordance = NA, N = NA, events = NA , std_beta = NA  # Add a column for the concordance index
)

for (j in 1:length(BA_measures)) {
  # Sex-adjusted model
  fit_gender <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_vas) ~", BA_measures[j], "+ gender")), data = to1po2_SveDem_D)
  coef_gender <- summary(fit_gender)
  row_index_gender <- (j * 2) - 1
  std_beta_gender <- calc_standardized_beta(fit_gender, to1po2_SveDem_D, BA_measures[j])
  HR_BA[row_index_gender, c(4, 5, 9, 12)] <- c(coef_gender$coefficients[1, c(1, 3, 5)], coef_gender$concordance[1])
  HR_BA[row_index_gender, c("HR", "CI_lower", "CI_upper")] <- coef_gender$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_gender, 10] <- cox.zph(fit_gender)$table[1, 3]
  HR_BA[row_index_gender, 11] <- cox.zph(fit_gender)$table["GLOBAL", "p"]
  HR_BA[row_index_gender, "N"] <- coef_gender$n
  HR_BA[row_index_gender, "std_beta"] <- std_beta_gender
  HR_BA[row_index_gender, "events"] <- fit_gender$nevent
  
  # Multivariable model
  fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_vas) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva   ")), data = to1po2_SveDem_D)
  coef_multivariable <- summary(fit_multivariable)
  row_index_multivariable <- j * 2
  std_beta_multivariable <- calc_standardized_beta(fit_multivariable, to1po2_SveDem_D, BA_measures[j])
  HR_BA[row_index_multivariable, c(4, 5, 9, 12)] <- c(coef_multivariable$coefficients[1, c(1, 3, 5)], coef_multivariable$concordance[1])
  HR_BA[row_index_multivariable, c("HR", "CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_multivariable, 10] <- cox.zph(fit_multivariable)$table[1, 3]
  HR_BA[row_index_multivariable, 11] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
  HR_BA[row_index_multivariable, "N"] <- coef_multivariable$n
  HR_BA[row_index_multivariable, "std_beta"] <- std_beta_multivariable
  HR_BA[row_index_multivariable, "events"] <- fit_multivariable$nevent
}

S_vas_results <- HR_BA

# Print the modified HR_BA_adrd data frame
print(HR_BA)
```

## 3.7 Lewy body dementia (SveDem)
```{r}
# BA measures to be tested
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Correct the initialization of HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),  # Repeat each BA measure twice
  outcome = rep("S_lew", times = length(BA_measures) * 2),
  model = rep(c("Sex-adjusted model", "Multivariable model"), times = length(BA_measures)),
  beta = NA, se = NA, HR = NA, CI_lower = NA, CI_upper = NA, p = NA, p_PH_BA = NA, global_p_PH_BA = NA,
  concordance = NA, N = NA, events = NA , std_beta = NA  # Add a column for the concordance index
)

for (j in 1:length(BA_measures)) {
  # Sex-adjusted model
  fit_gender <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_lew) ~", BA_measures[j], "+ gender")), data = to1po2_SveDem_D)
  coef_gender <- summary(fit_gender)
  row_index_gender <- (j * 2) - 1
  std_beta_gender <- calc_standardized_beta(fit_gender, to1po2_SveDem_D, BA_measures[j])
  HR_BA[row_index_gender, c(4, 5, 9, 12)] <- c(coef_gender$coefficients[1, c(1, 3, 5)], coef_gender$concordance[1])
  HR_BA[row_index_gender, c("HR", "CI_lower", "CI_upper")] <- coef_gender$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_gender, 10] <- cox.zph(fit_gender)$table[1, 3]
  HR_BA[row_index_gender, 11] <- cox.zph(fit_gender)$table["GLOBAL", "p"]
  HR_BA[row_index_gender, "N"] <- coef_gender$n
  HR_BA[row_index_gender, "std_beta"] <- std_beta_gender
  HR_BA[row_index_gender, "events"] <- fit_gender$nevent
  
  # Multivariable model
  fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_lew) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva   ")), data = to1po2_SveDem_D)
  coef_multivariable <- summary(fit_multivariable)
  row_index_multivariable <- j * 2
  std_beta_multivariable <- calc_standardized_beta(fit_multivariable, to1po2_SveDem_D, BA_measures[j])
  HR_BA[row_index_multivariable, c(4, 5, 9, 12)] <- c(coef_multivariable$coefficients[1, c(1, 3, 5)], coef_multivariable$concordance[1])
  HR_BA[row_index_multivariable, c("HR", "CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_multivariable, 10] <- cox.zph(fit_multivariable)$table[1, 3]
  HR_BA[row_index_multivariable, 11] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
  HR_BA[row_index_multivariable, "N"] <- coef_multivariable$n
  HR_BA[row_index_multivariable, "std_beta"] <- std_beta_multivariable
  HR_BA[row_index_multivariable, "events"] <- fit_multivariable$nevent
}

S_lew_results <- HR_BA

# Print the modified HR_BA_adrd data frame
print(HR_BA)
```

## 3.8 Frontotemporal dementia (SveDem)
```{r}
 # BA measures to be tested
 BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

 # Correct the initialization of HR_BA data frame
 HR_BA <- data.frame(
   BA = rep(BA_measures, each = 2),  # Repeat each BA measure twice
   outcome = rep("S_fro", times = length(BA_measures) * 2),
   model = rep(c("Sex-adjusted model", "Multivariable model"), times = length(BA_measures)),
   beta = NA, se = NA, HR = NA, CI_lower = NA, CI_upper = NA, p = NA, p_PH_BA = NA, global_p_PH_BA = NA,
   concordance = NA, N = NA, events = NA , std_beta = NA  # Add a column for the concordance index
 )

 for (j in 1:length(BA_measures)) {
   # Sex-adjusted model
   fit_gender <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_fro) ~", BA_measures[j], "+ gender")), data = to1po2_SveDem_D)
   coef_gender <- summary(fit_gender)
   row_index_gender <- (j * 2) - 1
   std_beta_gender <- calc_standardized_beta(fit_gender, to1po2_SveDem_D, BA_measures[j])
   HR_BA[row_index_gender, c(4, 5, 9, 12)] <- c(coef_gender$coefficients[1, c(1, 3, 5)], coef_gender$concordance[1])
   HR_BA[row_index_gender, c("HR", "CI_lower", "CI_upper")] <- coef_gender$conf.int[1, c(1, 3, 4)]
   HR_BA[row_index_gender, 10] <- cox.zph(fit_gender)$table[1, 3]
   HR_BA[row_index_gender, 11] <- cox.zph(fit_gender)$table["GLOBAL", "p"]
   HR_BA[row_index_gender, "N"] <- coef_gender$n
   HR_BA[row_index_gender, "std_beta"] <- std_beta_gender
   HR_BA[row_index_gender, "events"] <- fit_gender$nevent

   # Multivariable model
   fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_fro) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva   ")), data = to1po2_SveDem_D)
   coef_multivariable <- summary(fit_multivariable)
   row_index_multivariable <- j * 2
   std_beta_multivariable <- calc_standardized_beta(fit_multivariable, to1po2_SveDem_D, BA_measures[j])
   HR_BA[row_index_multivariable, c(4, 5, 9, 12)] <- c(coef_multivariable$coefficients[1, c(1, 3, 5)], coef_multivariable$concordance[1])
   HR_BA[row_index_multivariable, c("HR", "CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(1, 3, 4)]
   HR_BA[row_index_multivariable, 10] <- cox.zph(fit_multivariable)$table[1, 3]
   HR_BA[row_index_multivariable, 11] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
   HR_BA[row_index_multivariable, "N"] <- coef_multivariable$n
   HR_BA[row_index_multivariable, "std_beta"] <- std_beta_multivariable
   HR_BA[row_index_multivariable, "events"] <- fit_multivariable$nevent
 }

 S_fro_results <- HR_BA

 # Print the modified HR_BA_adrd data frame
 print(HR_BA)
```

## 3.9 Parkinsons dementia (SveDem)
```{r}
# BA measures to be tested
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Correct the initialization of HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),  # Repeat each BA measure twice
  outcome = rep("S_par", times = length(BA_measures) * 2),
  model = rep(c("Sex-adjusted model", "Multivariable model"), times = length(BA_measures)),
  beta = NA, se = NA, HR = NA, CI_lower = NA, CI_upper = NA, p = NA, p_PH_BA = NA, global_p_PH_BA = NA,
  concordance = NA, N = NA, events = NA , std_beta = NA  # Add a column for the concordance index
)

for (j in 1:length(BA_measures)) {
  # Sex-adjusted model
  fit_gender <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_par) ~", BA_measures[j], "+ gender")), data = to1po2_SveDem_D)
  coef_gender <- summary(fit_gender)
  row_index_gender <- (j * 2) - 1
  std_beta_gender <- calc_standardized_beta(fit_gender, to1po2_SveDem_D, BA_measures[j])
  HR_BA[row_index_gender, c(4, 5, 9, 12)] <- c(coef_gender$coefficients[1, c(1, 3, 5)], coef_gender$concordance[1])
  HR_BA[row_index_gender, c("HR", "CI_lower", "CI_upper")] <- coef_gender$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_gender, 10] <- cox.zph(fit_gender)$table[1, 3]
  HR_BA[row_index_gender, 11] <- cox.zph(fit_gender)$table["GLOBAL", "p"]
  HR_BA[row_index_gender, "N"] <- coef_gender$n
  HR_BA[row_index_gender, "std_beta"] <- std_beta_gender
  HR_BA[row_index_gender, "events"] <- fit_gender$nevent
  
  # Multivariable model
  fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_par) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva   ")), data = to1po2_SveDem_D)
  coef_multivariable <- summary(fit_multivariable)
  row_index_multivariable <- j * 2
  std_beta_multivariable <- calc_standardized_beta(fit_multivariable, to1po2_SveDem_D, BA_measures[j])
  HR_BA[row_index_multivariable, c(4, 5, 9, 12)] <- c(coef_multivariable$coefficients[1, c(1, 3, 5)], coef_multivariable$concordance[1])
  HR_BA[row_index_multivariable, c("HR", "CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_multivariable, 10] <- cox.zph(fit_multivariable)$table[1, 3]
  HR_BA[row_index_multivariable, 11] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
  HR_BA[row_index_multivariable, "N"] <- coef_multivariable$n
  HR_BA[row_index_multivariable, "std_beta"] <- std_beta_multivariable
  HR_BA[row_index_multivariable, "events"] <- fit_multivariable$nevent
}

S_par_results <- HR_BA

# Print the modified HR_BA_adrd data frame
print(HR_BA)
```

## 3.10 Unspecified dementia (SveDem)
```{r}
# BA measures to be tested
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Correct the initialization of HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),  # Repeat each BA measure twice
  outcome = rep("S_uns", times = length(BA_measures) * 2),
  model = rep(c("Sex-adjusted model", "Multivariable model"), times = length(BA_measures)),
  beta = NA, se = NA, HR = NA, CI_lower = NA, CI_upper = NA, p = NA, p_PH_BA = NA, global_p_PH_BA = NA,
  concordance = NA, N = NA, events = NA , std_beta = NA  # Add a column for the concordance index
)

for (j in 1:length(BA_measures)) {
  # Sex-adjusted model
  fit_gender <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_uns) ~", BA_measures[j], "+ gender")), data = to1po2_SveDem_D)
  coef_gender <- summary(fit_gender)
  row_index_gender <- (j * 2) - 1
  std_beta_gender <- calc_standardized_beta(fit_gender, to1po2_SveDem_D, BA_measures[j])
  HR_BA[row_index_gender, c(4, 5, 9, 12)] <- c(coef_gender$coefficients[1, c(1, 3, 5)], coef_gender$concordance[1])
  HR_BA[row_index_gender, c("HR", "CI_lower", "CI_upper")] <- coef_gender$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_gender, 10] <- cox.zph(fit_gender)$table[1, 3]
  HR_BA[row_index_gender, 11] <- cox.zph(fit_gender)$table["GLOBAL", "p"]
  HR_BA[row_index_gender, "N"] <- coef_gender$n
  HR_BA[row_index_gender, "std_beta"] <- std_beta_gender
  HR_BA[row_index_gender, "events"] <- fit_gender$nevent
  
  # Multivariable model
  fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_uns) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva   ")), data = to1po2_SveDem_D)
  coef_multivariable <- summary(fit_multivariable)
  row_index_multivariable <- j * 2
  std_beta_multivariable <- calc_standardized_beta(fit_multivariable, to1po2_SveDem_D, BA_measures[j])
  HR_BA[row_index_multivariable, c(4, 5, 9, 12)] <- c(coef_multivariable$coefficients[1, c(1, 3, 5)], coef_multivariable$concordance[1])
  HR_BA[row_index_multivariable, c("HR", "CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_multivariable, 10] <- cox.zph(fit_multivariable)$table[1, 3]
  HR_BA[row_index_multivariable, 11] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
  HR_BA[row_index_multivariable, "N"] <- coef_multivariable$n
  HR_BA[row_index_multivariable, "std_beta"] <- std_beta_multivariable
  HR_BA[row_index_multivariable, "events"] <- fit_multivariable$nevent
}

S_uns_results <- HR_BA

# Print the modified HR_BA_adrd data frame
print(HR_BA)
```

## 3.11 Other dementia (SveDem)
```{r}
# BA measures to be tested
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Correct the initialization of HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),  # Repeat each BA measure twice
  outcome = rep("S_oth", times = length(BA_measures) * 2),
  model = rep(c("Sex-adjusted model", "Multivariable model"), times = length(BA_measures)),
  beta = NA, se = NA, HR = NA, CI_lower = NA, CI_upper = NA, p = NA, p_PH_BA = NA, global_p_PH_BA = NA,
  concordance = NA, N = NA, events = NA , std_beta = NA  # Add a column for the concordance index
)

for (j in 1:length(BA_measures)) {
  # Sex-adjusted model
  fit_gender <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_oth) ~", BA_measures[j], "+ gender")), data = to1po2_SveDem_D)
  coef_gender <- summary(fit_gender)
  row_index_gender <- (j * 2) - 1
  std_beta_gender <- calc_standardized_beta(fit_gender, to1po2_SveDem_D, BA_measures[j])
  HR_BA[row_index_gender, c(4, 5, 9, 12)] <- c(coef_gender$coefficients[1, c(1, 3, 5)], coef_gender$concordance[1])
  HR_BA[row_index_gender, c("HR", "CI_lower", "CI_upper")] <- coef_gender$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_gender, 10] <- cox.zph(fit_gender)$table[1, 3]
  HR_BA[row_index_gender, 11] <- cox.zph(fit_gender)$table["GLOBAL", "p"]
  HR_BA[row_index_gender, "N"] <- coef_gender$n
  HR_BA[row_index_gender, "std_beta"] <- std_beta_gender
  HR_BA[row_index_gender, "events"] <- fit_gender$nevent
  
  # Multivariable model
  fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, S_oth) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva   ")), data = to1po2_SveDem_D)
  coef_multivariable <- summary(fit_multivariable)
  row_index_multivariable <- j * 2
  std_beta_multivariable <- calc_standardized_beta(fit_multivariable, to1po2_SveDem_D, BA_measures[j])
  HR_BA[row_index_multivariable, c(4, 5, 9, 12)] <- c(coef_multivariable$coefficients[1, c(1, 3, 5)], coef_multivariable$concordance[1])
  HR_BA[row_index_multivariable, c("HR", "CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(1, 3, 4)]
  HR_BA[row_index_multivariable, 10] <- cox.zph(fit_multivariable)$table[1, 3]
  HR_BA[row_index_multivariable, 11] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
  HR_BA[row_index_multivariable, "N"] <- coef_multivariable$n
  HR_BA[row_index_multivariable, "std_beta"] <- std_beta_multivariable
  HR_BA[row_index_multivariable, "events"] <- fit_multivariable$nevent
}

S_oth_results <- HR_BA

# Print the modified HR_BA_adrd data frame
print(HR_BA)
```

## 3.12 Binding and saving
```{r}
# Combine all data frames into one large data frame
combined_SveDem <- rbind(
  S_ADRD_results,
  S_oth_results,
  S_uns_results,
  S_par_results,
  S_fro_results,
  S_lew_results,
  S_vas_results,
  S_alz_vas_results,
  S_l_alz_results,
  S_e_alz_results
)
```

```{r}
write.table(as.data.frame(combined_SveDem), "combined_results_SveDem_main.csv", sep = ";",
            dec = ",",
            row.names = FALSE,
            quote = FALSE,
            fileEncoding = "Latin1")
```
