---
title: "CoxPH_subgroups_B"
output: html_notebook
author: Karolina Gustavsson
date: 2026-04-10
---

Input data: to1po2 from CoxPH.rmd
Covariates: sex, education level, income quartiles

# 1.0 Time to evenet analysis: age group

## 1.1 P-value function
```{r}
# Subgroup analysis (age cutoff 65)
p_value_results <- data.frame(
  Variable = character(),
  AgeGroup1 = character(),
  AgeGroup2 = character(),
  P_Value = numeric(),
  stringsAsFactors = FALSE
)
```

## 1.2 Median age = cut off 
```{r}
median(to1po2$age, na.rm = TRUE)
```

## 1.3 All (agegroup 45)
```{r}
BA_measures <- c("phenoage_advance_sd", "kdm_advance_sd") #"hd_log_res_sd"

# Initialize HR_BA_adrd data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),
  AgeGroup = rep(c("Under 45", "45 and over"), times = length(BA_measures)),
  model = rep("ADRD_agegroup_45", times = 2 * length(BA_measures)),
  beta = NA, se = NA, HR = NA,
  CI_lower = NA, CI_upper = NA, p = NA, P_PH_BA = NA, global_p_PH_BA = NA, concordance = NA, N = NA, events = NA, std_beta = NA) 

for (j in 1:length(BA_measures)) {
  # Loop over each BA measure
  for (age_group in c("Under 45", "45 and over")) {
    # Subset data based on age group
    if (age_group == "Under 45") {
      data_subset <- to1po2[to1po2$AgeAtFirstDate < 45, ]
    } else {
      data_subset <- to1po2[to1po2$AgeAtFirstDate >= 45, ]
    }

    # Fit Sex-adjusted model
    fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, ADRD) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva  ")), data = data_subset)
    coef_multivariable <- summary(fit_multivariable)
    std_beta <- calc_standardized_beta(fit_multivariable, data_subset, BA_measures[j])

    # Determine row index for updating HR_BA
    if (age_group == "Under 45") {
      row_index <- (j - 1) * 2 + 1
    } else {
      row_index <- j * 2
    }
    
    # Update HR_BA with model results
    HR_BA[row_index, c("beta", "se", "HR")] <- coef_multivariable$coefficients[1, c(1, 3, 2)]
    HR_BA[row_index, "p"] <- coef_multivariable$coefficients[1, 5]
    HR_BA[row_index, c("CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(3, 4)]
    HR_BA[row_index, "P_PH_BA"] <- cox.zph(fit_multivariable)$table[1, 3]
    HR_BA[row_index, "global_p_PH_BA"] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
    HR_BA[row_index, "concordance"] <- coef_multivariable$concordance[1]
    HR_BA[row_index, "N"] <- coef_multivariable$n
    HR_BA[row_index, "std_beta"] <- std_beta
    HR_BA[row_index, "events"] <- fit_multivariable$nevent
    #HR_BA[row_index_gender, "events"] <- coef_gender$nevent blir bara i en av dem
  }
}
ADRD_agegroup_45 <- HR_BA
# Print the updated HR_BA data frame
print(HR_BA)
```

## 1.4 Alzheimers (agegroup 45)
```{r}
BA_measures <- c("phenoage_advance_sd", "kdm_advance_sd") #"hd_log_res_sd"

# Initialize HR_BA_adrd data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),
  AgeGroup = rep(c("Under 45", "45 and over"), times = length(BA_measures)),
  model = rep("Alzheimers_agegroup_45", times = 2 * length(BA_measures)),
  beta = NA, se = NA, HR = NA,
  CI_lower = NA, CI_upper = NA, p = NA, P_PH_BA = NA, global_p_PH_BA = NA, concordance = NA, N = NA, events = NA, std_beta = NA ) #events =NA

for (j in 1:length(BA_measures)) {
  # Loop over each BA measure
  for (age_group in c("Under 45", "45 and over")) {
    # Subset data based on age group
    if (age_group == "Under 45") {
      data_subset <- to1po2[to1po2$AgeAtFirstDate < 45, ]
    } else {
      data_subset <- to1po2[to1po2$AgeAtFirstDate >= 45, ]
    }

    # Fit Sex-adjusted model
    fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, Alzheimers) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva  ")), data = data_subset)
    coef_multivariable <- summary(fit_multivariable)
    std_beta <- calc_standardized_beta(fit_multivariable, data_subset, BA_measures[j])

    # Determine row index for updating HR_BA
    if (age_group == "Under 45") {
      row_index <- (j - 1) * 2 + 1
    } else {
      row_index <- j * 2
    }
    
    # Update HR_BA with model results
    HR_BA[row_index, c("beta", "se", "HR")] <- coef_multivariable$coefficients[1, c(1, 3, 2)]
    HR_BA[row_index, "p"] <- coef_multivariable$coefficients[1, 5]
    HR_BA[row_index, c("CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(3, 4)]
    HR_BA[row_index, "P_PH_BA"] <- cox.zph(fit_multivariable)$table[1, 3]
    HR_BA[row_index, "global_p_PH_BA"] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
    HR_BA[row_index, "concordance"] <- coef_multivariable$concordance[1]
    HR_BA[row_index, "N"] <- coef_multivariable$n
    HR_BA[row_index, "std_beta"] <- std_beta
    HR_BA[row_index, "events"] <- fit_multivariable$nevent
    #HR_BA[row_index_gender, "events"] <- coef_gender$nevent blir bara i en av dem
  }
}
Alzh_agegroup_45 <- HR_BA
# Print the updated HR_BA data frame
print(HR_BA)
```

## 1.5 Vascular (agegroup 45)
```{r}
BA_measures <- c("phenoage_advance_sd", "kdm_advance_sd") #"hd_log_res_sd"

# Initialize HR_BA_adrd data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),
  AgeGroup = rep(c("Under 45", "45 and over"), times = length(BA_measures)),
  model = rep("Vascular_agegroup_45", times = 2 * length(BA_measures)),
  beta = NA, se = NA, HR = NA,
  CI_lower = NA, CI_upper = NA, p = NA, P_PH_BA = NA, global_p_PH_BA = NA, concordance = NA, N = NA, events = NA, std_beta = NA ) #events =NA

for (j in 1:length(BA_measures)) {
  # Loop over each BA measure
  for (age_group in c("Under 45", "45 and over")) {
    # Subset data based on age group
    if (age_group == "Under 45") {
      data_subset <- to1po2[to1po2$AgeAtFirstDate < 45, ]
    } else {
      data_subset <- to1po2[to1po2$AgeAtFirstDate >= 45, ]
    }

    # Fit Sex-adjusted model
    fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, Vascular) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva  ")), data = data_subset)
    coef_multivariable <- summary(fit_multivariable)
    std_beta <- calc_standardized_beta(fit_multivariable, data_subset, BA_measures[j])

    # Determine row index for updating HR_BA
    if (age_group == "Under 45") {
      row_index <- (j - 1) * 2 + 1
    } else {
      row_index <- j * 2
    }
    
    # Update HR_BA with model results
    HR_BA[row_index, c("beta", "se", "HR")] <- coef_multivariable$coefficients[1, c(1, 3, 2)]
    HR_BA[row_index, "p"] <- coef_multivariable$coefficients[1, 5]
    HR_BA[row_index, c("CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(3, 4)]
    HR_BA[row_index, "P_PH_BA"] <- cox.zph(fit_multivariable)$table[1, 3]
    HR_BA[row_index, "global_p_PH_BA"] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
    HR_BA[row_index, "concordance"] <- coef_multivariable$concordance[1]
    HR_BA[row_index, "N"] <- coef_multivariable$n
    HR_BA[row_index, "std_beta"] <- std_beta
    HR_BA[row_index, "events"] <- fit_multivariable$nevent
    #HR_BA[row_index_gender, "events"] <- coef_gender$nevent blir bara i en av dem
  }
}
Vasc_agegroup_45 <- HR_BA
# Print the updated HR_BA data frame
print(HR_BA)
```

## 1.6 Other (agegroup 45)
```{r}
BA_measures <- c("phenoage_advance_sd", "kdm_advance_sd") #"hd_log_res_sd"

# Initialize HR_BA_adrd data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),
  AgeGroup = rep(c("Under 45", "45 and over"), times = length(BA_measures)),
  model = rep("Other_agegroup_45", times = 2 * length(BA_measures)),
  beta = NA, se = NA, HR = NA,
  CI_lower = NA, CI_upper = NA, p = NA, P_PH_BA = NA, global_p_PH_BA = NA, concordance = NA, N = NA, events = NA, std_beta = NA ) #events =NA

for (j in 1:length(BA_measures)) {
  # Loop over each BA measure
  for (age_group in c("Under 45", "45 and over")) {
    # Subset data based on age group
    if (age_group == "Under 45") {
      data_subset <- to1po2[to1po2$AgeAtFirstDate < 45, ]
    } else {
      data_subset <- to1po2[to1po2$AgeAtFirstDate >= 45, ]
    }

    # Fit Sex-adjusted model
    fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, Other) ~", BA_measures[j], "+ gender + income_bins + NewUtbNiva  ")), data = data_subset)
    coef_multivariable <- summary(fit_multivariable)
    std_beta <- calc_standardized_beta(fit_multivariable, data_subset, BA_measures[j])

    # Determine row index for updating HR_BA
    if (age_group == "Under 45") {
      row_index <- (j - 1) * 2 + 1
    } else {
      row_index <- j * 2
    }
    
    # Update HR_BA with model results
    HR_BA[row_index, c("beta", "se", "HR")] <- coef_multivariable$coefficients[1, c(1, 3, 2)]
    HR_BA[row_index, "p"] <- coef_multivariable$coefficients[1, 5]
    HR_BA[row_index, c("CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(3, 4)]
    HR_BA[row_index, "P_PH_BA"] <- cox.zph(fit_multivariable)$table[1, 3]
    HR_BA[row_index, "global_p_PH_BA"] <- cox.zph(fit_multivariable)$table["GLOBAL", "p"]
    HR_BA[row_index, "concordance"] <- coef_multivariable$concordance[1]
    HR_BA[row_index, "N"] <- coef_multivariable$n
    HR_BA[row_index, "std_beta"] <- std_beta
    HR_BA[row_index, "events"] <- fit_multivariable$nevent
    #HR_BA[row_index_gender, "events"] <- coef_gender$nevent blir bara i en av dem
  }
}
Othe_agegroup_45 <- HR_BA
# Print the updated HR_BA data frame
print(HR_BA)
```

## 1.7 Binding
```{r}
combined_df_agegroup_45 <- rbind(
  ADRD_agegroup_45,
  Alzh_agegroup_45,
  Vasc_agegroup_45,
  Othe_agegroup_45
)
```

## 1.8 P-values comparing the two age groups
```{r}
# Create an empty data frame to store results
comparison_results <- data.frame()

# Loop through each unique BA and model combination
unique_combos <- unique(combined_df_agegroup_45[, c("BA", "model")])

for (i in seq_len(nrow(unique_combos))) {
  ba <- unique_combos$BA[i]
  model <- unique_combos$model[i]

  subset_rows <- combined_df_agegroup_45[combined_df_agegroup_45$BA == ba & 
                                           combined_df_agegroup_45$model == model, ]

  if (nrow(subset_rows) == 2 && all(c("Under 45", "45 and over") %in% subset_rows$AgeGroup)) {
    # Extract beta and SE
    beta_under45 <- subset_rows$beta[subset_rows$AgeGroup == "Under 45"]
    se_under45 <- subset_rows$se[subset_rows$AgeGroup == "Under 45"]

    beta_over45 <- subset_rows$beta[subset_rows$AgeGroup == "45 and over"]
    se_over45 <- subset_rows$se[subset_rows$AgeGroup == "45 and over"]

    # Z-test for difference in coefficients
    z <- (beta_under45 - beta_over45) / sqrt(se_under45^2 + se_over45^2)
    p_val <- 2 * (1 - pnorm(abs(z)))

    # Append to results
    comparison_results <- rbind(comparison_results, data.frame(
      BA = ba,
      model = model,
      AgeGroup1 = "Under 45",
      AgeGroup2 = "45 and over",
      p_value_group = p_val
    ))
  }
}

# Join the results back to your combined_df_agegroup_45 if needed
combined_with_p <- merge(combined_df_agegroup_45, comparison_results, by = c("BA", "model"), all.x = TRUE)

# Done! You can now export combined_with_p
```

## 1.9 Saving
```{r}
write.table(as.data.frame(combined_with_p), "combined_df_agegroup_45_main.csv", sep = ";",
            dec = ",",
            row.names = FALSE,
            quote = FALSE,
            fileEncoding = "Latin1")
```

# 2.0 Time to evenet analysis: sex

## 2.1 Making sex into a factor
```{r}
to1po2$new_gender <- to1po2$gender

to1po2$new_gender <- factor(to1po2$new_gender, levels = c("1", "2"), labels = c("Men", "Women"))
```

## 2.2 ADRD (sex)
```{r}
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Initialize HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),
  Gender = rep(c("Men", "Women"), times = length(BA_measures)),
  model = rep("ADRD_sex", times = 2 * length(BA_measures)),
  beta = NA, se = NA, HR = NA,
  CI_lower = NA, CI_upper = NA, p = NA, P_PH_BA = NA, global_p_PH_BA = NA, concordance = NA, N = NA, events = NA, std_beta = NA
)

for (j in 1:length(BA_measures)) {
  for (gender_group in c("Men", "Women")) {
    # Subset data based on new_sex
    data_subset <- to1po2[to1po2$new_gender == gender_group, ]

    # Fit Cox model for the BA measure (without 'new_gender' as a covariate)
    fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, ADRD) ~", BA_measures[j], "+ income_bins + NewUtbNiva  ")), data = data_subset)
    coef_multivariable <- summary(fit_multivariable)
    std_beta <- calc_standardized_beta(fit_multivariable, data_subset, BA_measures[j])

    # Determine row index for updating HR_BA
    row_index <- if (gender_group == "Men") (j - 1) * 2 + 1 else j * 2

    # Update HR_BA with model results
    HR_BA[row_index, c("beta", "se", "HR")] <- coef_multivariable$coefficients[1, c(1, 3, 2)]
    HR_BA[row_index, "p"] <- coef_multivariable$coefficients[1, 5]
    HR_BA[row_index, c("CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(3, 4)]
    HR_BA[row_index, "P_PH_BA"] <- cox.zph(fit_multivariable)$table[1, 3]
    HR_BA[row_index, "global_p_PH_BA"] <- cox.zph(fit_multivariable)$table[2, 3]
    HR_BA[row_index, "concordance"] <- coef_multivariable$concordance[1]
    HR_BA[row_index, "N"] <- coef_multivariable$n
    HR_BA[row_index, "std_beta"] <- std_beta
    HR_BA[row_index, "events"] <- fit_multivariable$nevent
  }
}
ADRD_sex <- HR_BA
# Print the updated HR_BA data frame
print(HR_BA)
```

## 2.3 Alzheimers (sex)
```{r}
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Initialize HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),
  Gender = rep(c("Men", "Women"), times = length(BA_measures)),
  model = rep("Alzheimers_sex", times = 2 * length(BA_measures)),
  beta = NA, se = NA, HR = NA,
  CI_lower = NA, CI_upper = NA, p = NA, P_PH_BA = NA, global_p_PH_BA = NA, concordance = NA, N = NA, events = NA, std_beta = NA
)

for (j in 1:length(BA_measures)) {
  for (gender_group in c("Men", "Women")) {
    # Subset data based on new_sex
    data_subset <- to1po2[to1po2$new_gender == gender_group, ]

    # Fit Cox model for the BA measure (without 'new_gender' as a covariate)
    fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, Alzheimers) ~", BA_measures[j], "+ income_bins + NewUtbNiva  ")), data = data_subset)
    coef_multivariable <- summary(fit_multivariable)
    std_beta <- calc_standardized_beta(fit_multivariable, data_subset, BA_measures[j])

    # Determine row index for updating HR_BA
    row_index <- if (gender_group == "Men") (j - 1) * 2 + 1 else j * 2

    # Update HR_BA with model results
    HR_BA[row_index, c("beta", "se", "HR")] <- coef_multivariable$coefficients[1, c(1, 3, 2)]
    HR_BA[row_index, "p"] <- coef_multivariable$coefficients[1, 5]
    HR_BA[row_index, c("CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(3, 4)]
    HR_BA[row_index, "P_PH_BA"] <- cox.zph(fit_multivariable)$table[1, 3]
    HR_BA[row_index, "global_p_PH_BA"] <- cox.zph(fit_multivariable)$table[2, 3]
    HR_BA[row_index, "concordance"] <- coef_multivariable$concordance[1]
    HR_BA[row_index, "N"] <- coef_multivariable$n
    HR_BA[row_index, "std_beta"] <- std_beta
    HR_BA[row_index, "events"] <- fit_multivariable$nevent
  }
}
Alzh_sex <- HR_BA
# Print the updated HR_BA data frame
print(HR_BA)
```

## 2.4 Vascular (sex)
```{r}
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Initialize HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),
  Gender = rep(c("Men", "Women"), times = length(BA_measures)),
  model = rep("Vascular_sex", times = 2 * length(BA_measures)),
  beta = NA, se = NA, HR = NA,
  CI_lower = NA, CI_upper = NA, p = NA, P_PH_BA = NA, global_p_PH_BA = NA, concordance = NA, N = NA, events = NA, std_beta = NA
)

for (j in 1:length(BA_measures)) {
  for (gender_group in c("Men", "Women")) {
    # Subset data based on new_sex
    data_subset <- to1po2[to1po2$new_gender == gender_group, ]

    # Fit Cox model for the BA measure (without 'new_gender' as a covariate)
    fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, Vascular) ~", BA_measures[j], "+ income_bins + NewUtbNiva  ")), data = data_subset)
    coef_multivariable <- summary(fit_multivariable)
    std_beta <- calc_standardized_beta(fit_multivariable, data_subset, BA_measures[j])

    # Determine row index for updating HR_BA
    row_index <- if (gender_group == "Men") (j - 1) * 2 + 1 else j * 2

    # Update HR_BA with model results
    HR_BA[row_index, c("beta", "se", "HR")] <- coef_multivariable$coefficients[1, c(1, 3, 2)]
    HR_BA[row_index, "p"] <- coef_multivariable$coefficients[1, 5]
    HR_BA[row_index, c("CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(3, 4)]
    HR_BA[row_index, "P_PH_BA"] <- cox.zph(fit_multivariable)$table[1, 3]
    HR_BA[row_index, "global_p_PH_BA"] <- cox.zph(fit_multivariable)$table[2, 3]
    HR_BA[row_index, "concordance"] <- coef_multivariable$concordance[1]
    HR_BA[row_index, "N"] <- coef_multivariable$n
    HR_BA[row_index, "std_beta"] <- std_beta
    HR_BA[row_index, "events"] <- fit_multivariable$nevent
  }
}
Vasc_sex <- HR_BA
# Print the updated HR_BA data frame
print(HR_BA)
```

## 2.5 Other (sex)
```{r}
BA_measures <- c("kdm_advance_sd", "phenoage_advance_sd")

# Initialize HR_BA data frame
HR_BA <- data.frame(
  BA = rep(BA_measures, each = 2),
  Gender = rep(c("Men", "Women"), times = length(BA_measures)),
  model = rep("Other_sex", times = 2 * length(BA_measures)),
  beta = NA, se = NA, HR = NA,
  CI_lower = NA, CI_upper = NA, p = NA, P_PH_BA = NA, global_p_PH_BA = NA, concordance = NA, N = NA, events = NA, std_beta = NA
)

for (j in 1:length(BA_measures)) {
  for (gender_group in c("Men", "Women")) {
    # Subset data based on new_sex
    data_subset <- to1po2[to1po2$new_gender == gender_group, ]

    # Fit Cox model for the BA measure (without 'new_gender' as a covariate)
    fit_multivariable <- coxph(as.formula(paste("Surv(AgeAtFirstDate, AgeAtLastDate, Other) ~", BA_measures[j], "+ income_bins + NewUtbNiva  ")), data = data_subset)
    coef_multivariable <- summary(fit_multivariable)
    std_beta <- calc_standardized_beta(fit_multivariable, data_subset, BA_measures[j])

    # Determine row index for updating HR_BA
    row_index <- if (gender_group == "Men") (j - 1) * 2 + 1 else j * 2

    # Update HR_BA with model results
    HR_BA[row_index, c("beta", "se", "HR")] <- coef_multivariable$coefficients[1, c(1, 3, 2)]
    HR_BA[row_index, "p"] <- coef_multivariable$coefficients[1, 5]
    HR_BA[row_index, c("CI_lower", "CI_upper")] <- coef_multivariable$conf.int[1, c(3, 4)]
    HR_BA[row_index, "P_PH_BA"] <- cox.zph(fit_multivariable)$table[1, 3]
    HR_BA[row_index, "global_p_PH_BA"] <- cox.zph(fit_multivariable)$table[2, 3]
    HR_BA[row_index, "concordance"] <- coef_multivariable$concordance[1]
    HR_BA[row_index, "N"] <- coef_multivariable$n
    HR_BA[row_index, "std_beta"] <- std_beta
    HR_BA[row_index, "events"] <- fit_multivariable$nevent
  }
}
Othe_sex <- HR_BA
# Print the updated HR_BA data frame
print(HR_BA)
```

## 2.6 Binding
```{r}
combined_df_sex <- rbind(
  ADRD_sex,
  Alzh_sex,
  Vasc_sex,
  Othe_sex
)
```

## 2.7 P-values comparing the sexes
```{r}
# Create an empty data frame to store results
gender_comparison_results <- data.frame()

# Loop through each unique BA and model combination
unique_combos <- unique(combined_df_sex[, c("BA", "model")])

for (i in seq_len(nrow(unique_combos))) {
  ba <- unique_combos$BA[i]
  model <- unique_combos$model[i]

  subset_rows <- combined_df_sex[combined_df_sex$BA == ba & 
                                  combined_df_sex$model == model, ]

  if (nrow(subset_rows) == 2 && all(c("Men", "Women") %in% subset_rows$Gender)) {
    # Extract beta and SE for each gender
    beta_men <- subset_rows$beta[subset_rows$Gender == "Men"]
    se_men <- subset_rows$se[subset_rows$Gender == "Men"]

    beta_women <- subset_rows$beta[subset_rows$Gender == "Women"]
    se_women <- subset_rows$se[subset_rows$Gender == "Women"]

    # Z-test for difference in coefficients
    z <- (beta_men - beta_women) / sqrt(se_men^2 + se_women^2)
    p_val <- 2 * (1 - pnorm(abs(z)))

    # Append result
    gender_comparison_results <- rbind(gender_comparison_results, data.frame(
      BA = ba,
      model = model,
      Gender1 = "Men",
      Gender2 = "Women",
      p_value_gender = p_val
    ))
  }
}

# Merge back to original dataframe if you want to include the p-values
combined_df_sex_with_p <- merge(combined_df_sex, gender_comparison_results, by = c("BA", "model"), all.x = TRUE)
```

## 2.8 Saving
```{r}
write.table(as.data.frame(combined_df_sex_with_p), "combined_sex_main.csv", sep = ";",
            dec = ",",
            row.names = FALSE,
            quote = FALSE,
            fileEncoding = "Latin1")
```
