---
title: "Table1_B"
output: html_notebook
author: Karolina Gustavsson
date: 2026-04-07
---
Before starting load CoxPH

# 1.0 Preparation

## 1.1 Load libraries
```{r}
options(repos = "http://nexus.ki.se/repository/cran.r-project.org")
.libPaths("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/shun_trial/lib")

library(haven)
library(broom)
library(dplyr)
library(flexsurv)
library(ggplot2)
library(htmlTable)
library(survival)
library(survey)
library(tidyr)
library(rmarkdown)
library(table1)
```

## 1.2 Add mortality status at end of follow up
```{r}
to1po2$Mortality <- ifelse(to1po2$status == 1, "Dead", "Alive")
label(to1po2$Mortality) <- "Mortality at follow up"

```

## 1.3 Add follow up time
```{r}
to1po2$FollowUpTime <- to1po2$AgeAtLastDate - to1po2$AgeAtFirstDate
label(to1po2$FollowUpTime) <- "Follow up time"
```

## 1.4 Recode sex
```{r}
# Create Sex column based on gender
to1po2$Sex <- ifelse(to1po2$gender == "1", "Men", 
                     ifelse(to1po2$gender == "2", "Women", NA))
```

## 1.5 Add missing to income quartiles
```{r}
# Define levels explicitly with "Missing" last
to1po2$Income_quartile <- factor(
  ifelse(is.na(to1po2$income_bins), "Missing", paste0("Q", to1po2$income_bins)),
  levels = c("Q1", "Q2", "Q3", "Q4", "Missing")
)

label(to1po2$Income_quartile) <- "Income quartiles"
```

## 1.6 Redcode education levels for more clarity
```{r} 
to1po2$Education_level <- ifelse(to1po2$NewUtbNiva %in% c(1, 2), "1 Basic",
                           ifelse(to1po2$NewUtbNiva %in% c(3, 4), "2. Secondary",
                           ifelse(to1po2$NewUtbNiva %in% c(5, 6, 7), "3. Post-secondary",
                           ifelse(is.na(to1po2$NewUtbNiva), "4. Unknown", "Other"))))

label(to1po2$Education_level) <- "Education level"
```

## 1.7 Recode fasting status
```{r}
to1po2$fasting_status <- ifelse(to1po2$fastestatus == "F", "Fasting overnight",
                          ifelse(to1po2$fastestatus == "N4", "Fasting 4h",
                          ifelse(to1po2$fastestatus %in% c("N", "S"), "Non-fasting",
                          "Unknown")))

label(to1po2$fasting_status) <- "Fasting status"
```

## 1.8 Adjust labels
```{r}
labels <- list(
  label(to1po2$AgeAtFirstDate) <- "Age at baseline (years, SD)",
  label(to1po2$AgeAtLastDate) <- "Age at follow up (years, SD)",
  label(to1po2$FollowUpTime) <- "Follow up time (years, SD)",
  label(to1po2$Mortality) <- "Mortality at follow up (N, %)",
  label(to1po2$Education_level) <- "Education level (N, %)",
  label(to1po2$Income_quartile) <- "Income quartiles (N, %)",
  label(to1po2$kdm) <- "KDMAge (years, SD)",
  label(to1po2$phenoage) <- "PhenoAge (years, SD)",
  label(to1po2$kdm_advance) <- "KDM Agegap (years, SD)",
  label(to1po2$phenoage_advance) <- "PhenoAge Agegap (years, SD)",
  label(to1po2$birthcohort) <- "Birth Cohort (N, %)",
  
  label(to1po2$S_Alb) <- "S-Albumin (g/L, SD)",
  label(to1po2$S_Krea) <- "S-Creatinine (umol/L, SD)",
  label(to1po2$TC) <- "S-Cholesterol (mmol/L, SD)",
  label(to1po2$TG) <- "S-Triglycerides (mmol/L, SD)",
  label(to1po2$S_K) <- "S-Potassium (mmol/L, SD)",
  label(to1po2$S_Urea) <- "S-Blood urea nitrogen (mmol/L, SD)",
  label(to1po2$S_LD) <- "S-Lactate dehydrogenase (ukat/L, SD)",
  label(to1po2$S_Ca) <- "S-Calcium (mmol/L, SD)",
  label(to1po2$S_Urat) <- "S-Uric acid (umol/L, SD)",
  label(to1po2$S_FAMN) <- "S-Fructosamine (mmol/L, SD)",
  label(to1po2$S_Hapt) <- "S-Haptoglobin (g/L, SD)",
  label(to1po2$fS_Gluk) <- "S-Glucose (mmol/L, SD)",
  
  label(to1po2$ADRD) <- "ADRD (N, %)",
  label(to1po2$Alzheimers) <- "Alzheimer's (N, %)",
  label(to1po2$Vascular) <- "Vascular (N, %)",
  label(to1po2$Other) <- "Other (N, %)",
  
  label(to1po2$S_ADRD) <- "SveDem ADRD (N, %)",
  label(to1po2$S_e_alz) <- "SveDem Early Alzheimer's (N, %)",
  label(to1po2$S_l_alz) <- "SveDem Late Alzheimer's (N, %)",
  label(to1po2$S_alz_vas) <- "SveDem Mixed (N, %)",
  label(to1po2$S_vas) <- "SveDem Vascular (N, %)",
  label(to1po2$S_fro) <- "SveDem Frontotemporal (N, %)",
  label(to1po2$S_lew) <- "SveDem Lewy Body (N, %)",
  label(to1po2$S_par) <- "SveDem Parkinson's (N, %)",
  label(to1po2$S_uns) <- "SveDem Unspecified (N, %)",
  label(to1po2$S_oth) <- "SveDem Other (N, %)")
```

## 1.9 Create two groups based on biological age (KDM)
```{r}
# Biologically younger and older
to1po2$kdm_a_group <- cut(
  to1po2$kdm_advance,
  breaks = quantile(to1po2$kdm_advance, probs = seq(0, 1, 0.5), na.rm = TRUE),
  include.lowest = TRUE,
  labels = c("Low Age Gap", "High Age Gap")
)

# Check result
table(to1po2$kdm_a_group)
```

## 1.10 Saving these groups for making the longitudinal table
```{r}
# Create dataframe with sampleID and KDM group
kdm_groups <- data.frame(
  sampleID = to1po2$sampleID,
  KDM_group = to1po2$kdm_a_group
)

# Save as CSV
write.table(
  kdm_groups,
  "table1_kdmgroups_baseline.csv",
  sep = ";",
  dec = ",",
  row.names = FALSE,
  quote = FALSE,
  fileEncoding = "Latin1"
)
```

## 1.11 Custom rendering function for table1
```{r}
render.custom <- function(x, ...) {
  if (is.numeric(x) && length(unique(na.omit(x))) == 2 && all(unique(na.omit(x)) %in% c(0, 1))) {
    # Binary variable: Render counts and percentages
    count <- sum(x, na.rm = TRUE)
    if (count == 0) {
      "-" # Display "-" if N is 0
    } else {
      sprintf("%d (%.2f%%)", count, 100 * mean(x, na.rm = TRUE)) # Show percent with two decimals
    }
  } else if (is.numeric(x)) {
    # Numeric variable: Render mean ± SD
    sprintf("%.2f ± %.2f", mean(x, na.rm = TRUE), sd(x, na.rm = TRUE))
  } else {
    # Fallback for other types
    table1::render.default(x, ...)
  }
}
```

## 1.12 Adding p-values
```{r}
pvalue <- function(x, ...) {
  ## Keep only the two comparison groups, exclude "Total"
  x <- x[1:2]
  
  y <- unlist(x)
  g <- factor(rep(seq_along(x), times = sapply(x, length)))
  
  keep <- !is.na(y) & !is.na(g)
  y <- y[keep]
  g <- g[keep]
  
  if (length(unique(g)) < 2) return("")
  
  ## Treat binary 0/1 as categorical
  is_binary <- is.numeric(y) && all(unique(y) %in% c(0, 1))
  
  if (is.numeric(y) && !is_binary) {
    p <- tryCatch(t.test(y ~ g)$p.value, error = function(e) NA)
  } else {
    tab <- table(y, g)
    p <- tryCatch({
      if (any(tab < 5)) {
        fisher.test(tab)$p.value
      } else {
        chisq.test(tab, correct = FALSE)$p.value
      }
    }, error = function(e) NA)
  }
  
  if (is.na(p)) {
    ""
  } else if (p < 0.001) {
    "<0.001"
  } else {
    sprintf("%.3f", p)
  }
}
```

## 2.0 Making tables

## 2.1 Making baseline characteristics table
```{r}
table_kdm_agegap_stratified_two <- table1(
  ~ Sex + AgeAtFirstDate + AgeAtLastDate + FollowUpTime +
    kdm + phenoage + kdm_advance + phenoage_advance + Mortality +
    Education_level + Income_quartile + birthcohort + fasting_status +
    S_Alb + S_Krea + TC + TG + S_K + S_Urea +
    S_LD + S_Ca + S_Urat + S_FAMN +
    S_Hapt + fS_Gluk + ADRD + Alzheimers + Vascular + Other + 
    S_ADRD + S_e_alz + S_l_alz + S_alz_vas +S_vas + S_fro + S_par + S_lew + S_oth + S_uns | kdm_a_group,
  data = to1po2,
  overall = "Total",
  render = render.custom,
  label = labels,
  extra.col = list(`P-value` = pvalue) #optional
  )

print(table_kdm_agegap_stratified_two)
```

```{r}
write.table(as.data.frame(table_kdm_agegap_stratified_two), "table_kdm_agegap_stratified_two_p.csv", sep = ";",
            dec = ",",
            row.names = FALSE,
            quote = FALSE,
            fileEncoding = "Latin1")
```

## 2.2 Making table with follow up data
```{r}
followuptime_ADRD_b <- table1(
  ~ AgeAtFirstDate + AgeAtLastDate + FollowUpTime | ADRD,
  data = to1po2,
  overall = "Total",
  render = render.custom,
  label = labels,
  extra.col = list(`P-value` = pvalue) #optional
  )

print(followuptime_ADRD_b)
```

```{r}
write.table(as.data.frame(followuptime_ADRD_b), "followuptime_ADRD_b_p.csv", sep = ";",
            dec = ",",
            row.names = FALSE,
            quote = FALSE,
            fileEncoding = "Latin1")
```
